---
title: "Case selection"
author: "Anthony Staines & Anne O'Farrell"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
    latex_engine: xelatex
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
    df_print: tibble
---

# Identify cases of TBI for further analysis

```{r setup, include=FALSE}
rm(list = ls())
library(lubridate)
library(gt)
library(arsenal)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE,cache = TRUE, message= FALSE, results = 'hide')
```

# Read in data

```{r}
TBI_w <- readRDS('data/TBI_w.Rds')
Dictionary <- readRDS('data/Dictionary.Rds')

```

Order the dictionary correctly, and add the Row number as a variable

```{r}

Dictionary <- Dictionary %>%
  arrange(ICD_code) %>%
  mutate(Row = row_number())

```

# Plan

We start with `r length(TBI_w$ID) ` cases available for analysis. These are all hospital discharges with codes in S00 to S99 from 2013 to late 2020.

The aim is to reduce these to the TBI cases, but there is no specific TBI code in ICD10, so we follow an algorithm from previous work in New South Wales - 

Pozzato I, Tate RL, Rosenkoetter U, Cameron ID. Epidemiology of hospitalised traumatic brain injury in the state of New South Wales, Australia: a population-based study. Australian and New Zealand Journal of Public Health [Internet]. 2019 [cited 2021 Feb 2];43(4):382–8. Available from: https://onlinelibrary.wiley.com/doi/abs/10.1111/1753-6405.12878



# Record selection

The primary selection was that all records with any code in S00 to S99 were included.

Focussing further on these, and following Pazzato et al. we focus on people with a skull fracture, a recorded intracranial injury, or a crush injury to the head. All cases with one or more of these codes, where an external cause of injury was coded were included. This gives us `r TBI_w %>% filter((Skull | IC_injury) & External)  %>% summarize(N=n()) #25958` cases immediately eligible.

```{r, results = 'asis'}
T1 <- tableby(External ~ Skull + IC_injury, data=TBI_w, test=FALSE, total=FALSE)

 summary(T1,text=TRUE)
```

Breaking these down more explicitly we get this :-

```{r, results = 'show'}
TBI_w %>% filter(Skull )  %>% summarize(N=n()) # 6676
TBI_w %>% filter(IC_injury)  %>% summarize(N=n()) # 25154

TBI_w %>% filter(Skull & IC_injury)  %>% summarize(N=n()) # 3979
TBI_w %>% filter(Skull | IC_injury)  %>% summarize(N=n()) # 27851

TBI_w %>% filter((Skull | IC_injury) & External)  %>% summarize(N=n()) # 25958 These are our key group 

TBI_w %>% filter((Skull) & !External)  %>% summarize(N=n()) # 340 These have no external cause identified.
TBI_w %>% filter((IC_injury) & !External)  %>% summarize(N=n()) # 1730 These also have no external cause identified.
TBI_w %>% filter((Skull & IC_injury) & !External)  %>% summarize(N=n()) # 177 These have no external cause identified.
TBI_w %>% filter((Skull | IC_injury) & !External)  %>% summarize(N=n()) # 1893 These have no external cause identified.
```

There are a number of sets of cases for whom it is appropriate to carry out a manual review. The aim of this review is, within the limits imposed by the available data, to identify possible additional TBI cases. When reviewing all classifications of cases were reviewed, and all the recorded ICD codes were inspected.

## Check 1 Loss of consciousness and/or post-traumatic amnesia.

The first group for manual review are those with LOC/PTA recorded, but no External cause given.


```{r}
TBI_w %>% filter((LOC | PTA))  %>% summarize(N=n()) # 12106 with LOC/PTA
TBI_w %>% filter((LOC | PTA) & External)  %>% summarize(N=n()) # 11091 with LOC/PTA
TBI_w %>% filter((LOC | PTA) & !External)  %>% summarize(N=n()) # 1015 with LOC/PTA and no external
```

There are a total of `r TBI_w %>% filter((LOC | PTA))  %>% summarize(N=n()) # 12109` cases with loss of consciousness and/or PTA recorded. Of these `r TBI_w %>% filter((LOC | PTA) & External)  %>% summarize(N=n()) # 11094` have an external cause recorded and only `r TBI_w %>% filter((LOC | PTA) & !External)  %>% summarize(N=n()) # 1015` do not.


```{r}
TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury))  %>% summarize(N=n()) # 11976 with a reported skull or intracranial injury and LOC/PTA

TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury))  %>% filter(External)  %>% summarize(N=n()) # These 10964 have an external cause identified.
TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury)) %>% filter(!External)  %>% summarize(N=n()) # These 1012 have no external cause identified.

```

Of those with LOC/PTA recorded, the large majority, `r TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury))  %>% summarize(N=n()) # 11979` also have a reported skull or intra-cranial injury. Most of these, `r TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury))  %>% filter(External)  %>% summarize(N=n()) #10967` have an external cause coded. Only `r TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury)) %>% filter(!External)  %>% summarize(N=n()) #1012` do not.

```{r}
TBI_w %>% filter((LOC | PTA) & External)  %>% summarize(N=n()) # 11091 External cause given
TBI_w %>% filter((LOC | PTA) & !External)  %>% summarize(N=n()) # 1015 No external cause given

TBI_w %>% filter((LOC | PTA) & !External & Undetermined)  %>% summarize(N=n()) # 16 Undetermined intent
TBI_w %>% filter((LOC | PTA) & !External & Assaults)  %>% summarize(N=n()) # 0 Assaults
TBI_w %>% filter((LOC | PTA) & !External & Selfharm)  %>% summarize(N=n()) # 46 Self harm
TBI_w %>% filter((LOC | PTA) & !External & Syncope)  %>% summarize(N=n()) # 352 Syncope

TBI_w %>% filter((LOC | PTA) & !External & !Selfharm)  %>% summarize(N=n()) # 969
TBI_w %>% filter((LOC | PTA) & !External & !Selfharm & !Undetermined)  %>% summarize(N=n()) # 934
TBI_w %>% filter((LOC | PTA) & !External & !Selfharm & !Undetermined & !Syncope)  %>% summarize(N=n()) # 591

TBI_w %>% filter((LOC | PTA) & !External & !Selfharm & !Undetermined & !Syncope & !(Skull | IC_injury))  %>% summarize(N=n()) # 3
```

Of the `r TBI_w %>% filter((LOC | PTA)) %>% filter((Skull | IC_injury)) %>% filter(!External)  %>% summarize(N=n()) #1012` with no external cause given, these can be further broken down. Only `r TBI_w %>% filter((LOC | PTA) & !External & Undetermined)  %>% summarize(N=n()) # 16` are recorded as undetermined intent, while `r TBI_w %>% filter((LOC | PTA) & !External & Selfharm)  %>% summarize(N=n()) # 46` are recorded as self harm. Quite a lot of these cases `r TBI_w %>% filter((LOC | PTA) & !External & Syncope)  %>% summarize(N=n()) # 352 ` are coded as syncope. These three groups were not considered further.

Of the remainder, only `r TBI_w %>% filter((LOC | PTA) & !External & !Selfharm & !Undetermined & !Syncope & !(Skull | IC_injury))  %>% summarize(N=n()) # 3` were not also coded as Skull fracture or intra-cranial injury. Accordingly all `r TBI_w %>% filter((LOC | PTA) & !External & !Selfharm & !Undetermined & !Syncope)  %>% summarize(N=n()) # 609` were separated out for manual review.


```{r}
# A check file is prepared
# These have LOC/PTA but no external cause given, and no indication of intent or syncope
ToBeChecked1 <- TBI_w %>% filter((LOC | PTA) & !External & !Selfharm & !Undetermined & ! Syncope)  %>% #summarize(N=n())
  select(ID:PTAn,Dx_1_name_CCS_IM, Dx_1_group_CCS_IM, Proc_1_ACHI_code, Proc_1_ACHI_name, Proc_1_group_RCs, Proc_1_name_RCs, DRG_name, DRG_code, ICD_name_Dx_1:ICD_name_Dx_30) # 625 cases

ToBeChecked1 %>% filter(Proc_1_name_RCs == 'No proc 1') %>% summarize(N= n()) # 352 had no primary procedure
ToBeChecked1 %>% filter(Proc_1_name_RCs != 'No proc 1') %>% summarize(N= n()) # 273 had a procedure


write_csv(ToBeChecked1,'data/ToBeChecked1.csv')
```

These cases covered a wide range of procedure codes.

```{r, results = 'asis'}

pander::pander(ToBeChecked1 %>% group_by(Proc_1_group_RCs) %>% summarize(N= n()) %>% arrange(desc(N)))
```

#Read back in the manually amended files.

These `r length(ToBeChecked1$ID) #591` cases, where there is no external code identified, but there is either LOC, or PTA, or both were to be manually reviewed to identify missed TBI cases.

```{r}
ToBeCheckedRevised1 <- readxl::read_xlsx('data/ToBeCheckedRevised1.xlsx',
                                         sheet = 'ToBeCheckedRevised1') # 591
glimpse(ToBeCheckedRevised1)
Note1 <- ToBeCheckedRevised1 %>% select(ID,E_MRN, Manual_check, Note, Detail, External_detail) %>%
  mutate(Detail_lump = fct_lump_n(Detail,6))

Note1 %>%
  anti_join(TBI_w,  by = c("E_MRN"))
TBI_w <- TBI_w %>%
  left_join(Note1,  by = c("ID", "E_MRN"))

```

For each case reviewed, a decision was made as to whether it fitted in the category of Trauma, the more specific category of Brain Injury, some other identifiable likely cause of loss-of-consciousness (Other), or whether it could not be reasonably determined, based on the data provided, what this was (Unknown). Cases identified as brain injury were further classified as TBI or Other. For example a discharge coded as Drowning but with a recorded cerebral contusion was coded as not a TBI. Cases identified as Other were further coded is there was a likely cause coded, which could explain the loss-of-consciousness. Examples included sepsis, seizures, or a cardiac arrythmia. For cases identified as possible TBI or Trauma, all the codes given were inspected to see if any indication of an external cause was given. The commonest of these was the location of the injury, and these were recorded.


```{r, results = 'asis'}
pander::pander(table(Note1$Note))
```

On manual checking they fell into 4 groups, as shown in the table. A small number had codes strongly suggesting traumatic brain injury, but had no identified external cause given. The largest group had no identifiable cause coded for the loss of consciousness. 

```{r, results = 'asis'}
pander::pander(table(Note1$Detail_lump))
```

Many had a code recorded, for example, epilepsy, cardiac arrythmia, stroke, or acute MI, which seemed likely to explain their LOC. A proportion had had other trauma, often a purely facial trauma, a nasal fracture, or a fractured jaw or tooth.

```{r, results = 'asis'}
pander::pander(table(Note1$External_detail))
```

Very few had any component of external causes recorded. In some cases there was an indication of the location or the setting for the event.

## Check 2 Skull fracture or Intracranial injury recorded

The second set of cases for further consideration as to eligibility, or otherwise, are those where there is no external code, but where either a Skull fracture, an intra-cranial injury, or both is recorded.


```{r}
TBI_w %>% filter(Skull) %>% summarize(N=n()) # 6676
TBI_w %>% filter(IC_injury) %>% summarize(N=n()) # 25154
TBI_w %>% filter((Skull & IC_injury))  %>% summarize(N=n()) # 3979
TBI_w %>% filter((Skull | IC_injury))  %>% summarize(N=n()) # 27851

TBI_w %>% filter(Skull & !External) %>% summarize(N=n()) # 340
TBI_w %>% filter(IC_injury & !External) %>% summarize(N=n()) # 1730

TBI_w %>% filter((Skull | IC_injury) & External)  %>% summarize(N=n()) # 25958
TBI_w %>% filter((Skull | IC_injury) & !External)  %>% summarize(N=n()) # 1893
```

There were `r TBI_w %>% filter(Skull) %>% summarize(N=n()) # 6676` records with a skull fracture recorded, and `r TBI_w %>% filter(IC_injury) %>% summarize(N=n()) # 25154` with an intracranial injury recorded. A total of `r TBI_w %>% filter((Skull | IC_injury))  %>% summarize(N=n()) # 27851` had one or other or both of these codes.

Of these the large majority `r TBI_w %>% filter((Skull | IC_injury) & External)  %>% summarize(N=n()) # 25958` had an external cause of injury coded, and only `r TBI_w %>% filter((Skull | IC_injury) & !External)  %>% summarize(N=n()) # 1893` did not.


```{r}
TBI_w %>% filter((Skull | IC_injury) & !External & Selfharm)  %>% summarize(N=n()) # 166
TBI_w %>% filter((Skull | IC_injury) & !External & Undetermined)  %>% summarize(N=n()) # 66
TBI_w %>% filter((Skull | IC_injury) & !External & Assaults)  %>% summarize(N=n()) # 0
TBI_w %>% filter((Skull | IC_injury) & !External & Syncope)  %>% summarize(N=n()) # 377


TBI_w %>% filter((Skull | IC_injury) & !External & !is.na(Note))  %>% summarize(N=n()) # 588
TBI_w %>% filter((Skull | IC_injury) & !External & is.na(Note))  %>% summarize(N=n()) # 1305
```

Of these, `r TBI_w %>% filter((Skull | IC_injury) & !External & Selfharm)  %>% summarize(N=n()) # 166` were coded as self harm, only `r TBI_w %>% filter((Skull | IC_injury) & !External & Undetermined)  %>% summarize(N=n()) # 39` were coded as undetermined, and `r TBI_w %>% filter((Skull | IC_injury) & !External & Syncope)  %>% summarize(N=n()) # 377` were coded as syncope.
`
There are `r TBI_w %>% filter((Skull | IC_injury) & !External & !is.na(Note))  %>% summarize(N=n())` cases with codes which have already been manually checked, as described in the previous section.

There are `r TBI_w %>% filter((Skull | IC_injury) & !External & !Selfharm & is.na(Note))  %>% summarize(N=n())` cases with codes which have not already been manually checked as outlined in the previous section.

```{r}
TBI_w %>% filter((Skull | IC_injury) & !External & !Assaults & !Selfharm & !Syncope & is.na(Note)) %>% summarise(N=n()) # 765
TBI_w %>% filter((Skull | IC_injury) & !External & !Assaults & !Selfharm & !Syncope) %>% summarise(N=n()) # 1353

TBI_w %>% filter((Skull | IC_injury) & !External & is.na(Note)) %>% summarise(N=n()) # 1305
TBI_w %>% filter((Skull | IC_injury) & !External & !is.na(Note)) %>% summarise(N=n()) # 588

```

```{r}

# A check file is prepared
ToBeChecked2 <- TBI_w %>% filter((Skull | IC_injury) & !External & !Assaults & !Selfharm & !Syncope & is.na(Note))  %>%
  select(ID:PTAn,Dx_1_name_CCS_IM, Dx_1_group_CCS_IM, Proc_1_ACHI_code, Proc_1_ACHI_name, Proc_1_group_RCs, Proc_1_name_RCs, Manual_check, Note, Detail, External, DRG_name, DRG_code, ICD_name_Dx_1:ICD_name_Dx_30, ICD_code_Dx_1:ICD_code_Dx_30) # 765

write_csv(ToBeChecked2,'data/ToBeChecked2.csv')

ToBeChecked2 %>% select(Dx_1_group_CCS_IM, Dx_1_name_CCS_IM) %>% group_by(Dx_1_group_CCS_IM, Dx_1_name_CCS_IM) %>% summarize(N = n()) %>% arrange(desc(N))
ToBeChecked2 %>% select(DRG_code, DRG_name) %>% group_by(DRG_code, DRG_name) %>% summarize(N = n()) %>% arrange(desc(N))
ToBeChecked2 %>% select(Proc_1_ACHI_code, Proc_1_ACHI_name) %>% group_by(Proc_1_ACHI_code, Proc_1_ACHI_name) %>% summarize(N = n()) %>% arrange(desc(N))

# Extensive manual checking suggests the following
# Leave out if CCS_IM group not Injury & Poisoning
# Leave out if CCS_IM name not Fracture skull & face or Intracranial injury
#

ToBeChecked2 %>%
  select(Dx_1_name_CCS_IM, Dx_1_group_CCS_IM) %>%
  group_by(Dx_1_name_CCS_IM, Dx_1_group_CCS_IM) %>%
  filter(Dx_1_group_CCS_IM == 'Injury & poisoning') %>%
  filter(Dx_1_name_CCS_IM == 'Fracture skull & face' | Dx_1_name_CCS_IM == 'Intracranial injury') %>%
  summarise(N = n()) %>% arrange( desc(N))

ToBeChecked2 %>%
  select(Dx_1_name_CCS_IM, Dx_1_group_CCS_IM) %>%
  group_by(Dx_1_name_CCS_IM, Dx_1_group_CCS_IM) %>%
  filter(Dx_1_group_CCS_IM == 'Injury & poisoning') %>%
  filter(Dx_1_name_CCS_IM == 'Fracture skull & face' | Dx_1_name_CCS_IM == 'Intracranial injury') %>%
  summarise(N = n()) %>% arrange( desc(N))

# Further manual checking suggests that cases can be discarded if
#  their Clinical classification group is not 'Injury & Poisoning', and their subgroup is not either 
#  'Fracture skull & face' or 'Intracranial injury')
# Cases where care was provided by other speciality groups were reviewed manually, and those grouped as
# 'Cardiology', 'Haematology', 'Endoscope', 'Not mapped proc 1', 'Anaesthetic & pain relief', 'Maxilliofacial & dental',
# 'Respiratory medicine',  'Otolaryngology', 'Urology ', 'Oncology', 'Vascular', 'Plastic surgery',
# 'Trauma orthopaedic other surgery', and 'Imaging and testing' were omitted.
# Finally remaining questionable cases where no LOC was recorded were omitted.
# All cases reporting post-traumatic amnesia were included.
#
ToBeChecked2 %>% # 1629
  ungroup() %>%
  filter(Dx_1_group_CCS_IM == 'Injury & poisoning') %>% # 977
  filter(Dx_1_name_CCS_IM == 'Fracture skull & face' | Dx_1_name_CCS_IM == 'Intracranial injury') %>% # 864
  filter(Proc_1_group_RCs %in% c('No proc 1', 'Medical interventions', 'Allied health interventions',
                                 'Neurosurgery','Paediatric medicine', 'Non surgery')) %>%  # 741
# Omit the following c('Cardiology', 'Haematology', 'Endoscope', 'Not mapped proc 1', 'Anaesthetic & pain relief',
#                       'Maxilliofacial & dental', 'Respiratory medicine',  'Otolaryngology', 'Urology ', 'Oncology',
#                       'Vascular', 'Plastic surgery', 'Trauma orthopaedic other surgery', 'Imaging and testing')
  filter(DRG_code != 'B80Z' & DRG_code != 'D04A') %>% # 336 Other loss of consciousness, Maxillary surgery
  filter(LOC | PTA) %>% #27
  select(Dx_1_name_CCS_IM, Dx_1_group_CCS_IM, DRG_code)

```

There are a total of  `r TBI_w %>% filter((Skull | IC_injury) & !External & !Selfharm & !Syncope & is.na(Note))  %>% summarize(N=n()) # 765` cases which required further manual examination. These are cases recorded as having a skull fracture or an intracranial injury, who have no external code, no indication of self-harm, assault, syncope, and which were not manually checked when cases with LOC were checked.

```{r}
ToBeCheckedRevised2 <- readxl::read_xlsx('data/ToBeCheckedRevised2.xlsx',
                                         sheet = 'ToBeCheckedRevised2')
Note2 <- ToBeCheckedRevised2 %>% select(ID,E_MRN, Manual_check, Note, Detail, External_detail) %>%
  mutate(Detail_lump = fct_lump_n(Detail,8))

Note2 %>%
  anti_join(TBI_w,  by = c("E_MRN"))
TBI_w <- TBI_w %>%
  left_join(Note2,  by = c("ID", "E_MRN"))
```

No cases occurred in both lists for manual checking. The two sets of notes, and cases details added manually need to be merged into one.

```{r}
# Tidy up variables
## Check - should all be zero - no records with two sets of notes etc.
TBI_w %>%
  filter(!is.na(Manual_check.x) & !is.na(Manual_check.y)) %>%
  summarise(n())
TBI_w %>%
  filter(!is.na(Note.x) & !is.na(Note.y)) %>%
  summarise(n())
TBI_w %>%
  filter(!is.na(Detail.x) & !is.na(Detail.y)) %>%
  summarise(n())
TBI_w %>%
  filter(!is.na(External_detail.x) & !is.na(External_detail.y)) %>%
  summarise(n())
TBI_w %>%
  filter(!is.na(Detail_lump.x) & !is.na(Detail_lump.y)) %>%
  summarise(n())

# Combine - many values are missing, so need to be reset to '' for combining.
## check it works
TBI_w %>%
  mutate(Manual_check = str_c(str_replace_na(Manual_check.x,''),
                      str_replace_na(Manual_check.y,''))) %>%
  mutate(Manual_check = ifelse(Manual_check == '',NA,Manual_check)) %>%
  filter(!is.na(Manual_check)) %>%
  select(Manual_check) %>%
  distinct()

TBI_w %>%
  mutate(Note = str_c(str_replace_na(Note.x,''),
                      str_replace_na(Note.y,''))) %>%
  mutate(Note = ifelse(Note == '',NA,Note)) %>%
  filter(!is.na(Note)) %>%
  select(Note) %>%
  distinct()

TBI_w %>%
  mutate(Detail = str_c(str_replace_na(Detail.x,''),
                      str_replace_na(Detail.y,''))) %>%
  mutate(Detail = ifelse(Detail == '',NA,Detail)) %>%
  filter(!is.na(Detail)) %>%
  select(Detail) %>%
  distinct()

TBI_w %>%
  mutate(Detail_lump = str_c(str_replace_na(Detail_lump.x,''),
                      str_replace_na(Detail_lump.y,''))) %>%
  mutate(Detail_lump = ifelse(Detail_lump == '',NA,Detail_lump)) %>%
  filter(!is.na(Detail_lump)) %>%
  select(Detail_lump) %>%
  distinct()

TBI_w %>%
  mutate(External_detail = str_c(str_replace_na(External_detail.x,''),
                      str_replace_na(External_detail.y,''))) %>%
  mutate(External_detail = ifelse(External_detail == '',NA,External_detail)) %>%
  filter(!is.na(External_detail)) %>%
  select(External_detail) %>%
  distinct()

# Do it
TBI_w <- TBI_w %>%
  mutate(Manual_check = str_c(str_replace_na(Manual_check.x,''),
                      str_replace_na(Manual_check.y,''))) %>%
  mutate(Manual_check = ifelse(Manual_check == '',NA,Manual_check)) %>%
  mutate(Manual_check = as.numeric(Manual_check)) %>%
  mutate(Note = str_c(str_replace_na(Note.x,''),
                      str_replace_na(Note.y,''))) %>%
  mutate(Note = ifelse(Note == '',NA,Note)) %>%
  mutate(Detail = str_c(str_replace_na(Detail.x,''),
                      str_replace_na(Detail.y,''))) %>%
  mutate(Detail = ifelse(Detail == '',NA,Detail)) %>%
  mutate(Detail_lump = str_c(str_replace_na(Detail_lump.x,''),
                      str_replace_na(Detail_lump.y,''))) %>%
  mutate(Detail_lump = ifelse(Detail_lump == '',NA,Detail_lump)) %>%
  mutate(External_detail = str_c(str_replace_na(External_detail.x,''),
                      str_replace_na(External_detail.y,''))) %>%
  mutate(External_detail = ifelse(External_detail == '',NA,External_detail)) %>%
  select(-(ends_with('.x') | ends_with('.y')))
  
glimpse(TBI_w)
table(TBI_w$Manual_check)
```

Of the cases derived from manual checking a small number are likely to be traumatic brain injuries, but have had no valid external cause codes provided. For a small number codes are given which suggest a location or a mechanism of injury, but no intent was identified.


```{r, results = 'asis'}
pander::pander(table(Note1$Note),caption='Main groups of additional cases - Check 1. LOC')
pander::pander(table(Note2$Note),caption='Main groups of additional cases - Check 2 Skull fracture or intracranial injury recorded.')
```

Manual checking suggests that very few of those with an LOC recorded were likely to have had traumatic brain injury, but that many of those with a recorded skull fracture or intracranial injury, were likely to be traumatic brain injuries.


```{r, results = 'asis'}
pander::pander(table(Note1$Detail_lump),caption='Main clinical details coded - Check 1. LOC')
pander::pander(table(Note2$Detail_lump),caption='Main clinical details coded - Check 2 Skull fracture or intracranial injury recorded.')
```

The first group (LOC and/or PTA) added few cases to the list of possible TBI's, but the second group (Skull fracture or Intra-cranial injury) added many more, as might be expected.  Cases included were those with codes consistent with injury, but with no other likely explanation for their fracture or intra-cranial injury recorded in the discharge codes.

```{r, results = 'asis'}
pander::pander(table(Note1$External_detail),caption='Main external codes recorded - Check 1. LOC')
pander::pander(table(Note2$External_detail),caption='Main external codes recorded - Check 2 Skull fracture or intracranial injury recorded.')
```

Although there were more external codes recorded in the second set of records checked, than in the first, 'Unknown' was still very prevalent.

```{r}
TBI_w %>% filter(((Skull | IC_injury | PTA) & External)) %>% summarise(N=n()) # 26085
TBI_w %>% filter(((Skull | IC_injury | PTA) & !External)) %>% summarise(N=n()) # 1896

TBI_w %>% filter(Detail == 'TBI') %>% summarise(N=n()) # 321
TBI_w %>% filter(((Skull | IC_injury | PTA) & External) | Detail == 'TBI') %>% summarise(N=n()) # 26406

CIs <- round(100*Hmisc::binconf(321,26406),2)

100-CIs

#  26085 321
#  
#
#

TestTable <- as.table(matrix(c(26085,70438,321,1575),nrow=2,byrow=TRUE))
epiR::epi.tests(TestTable)

```

Overall another `r TBI_w %>% filter(!is.na(Note)) %>% summarize(N=n()) # 1356` records were manually reviewed. Of these `r TBI_w %>% filter(!is.na(Note) & Detail == 'TBI') %>% summarize(N=n()) # 321` were probably potentially eligible to be recorded as TBI cases. It's important to emphasise that there was almost no information about intent, or mechanism recorded in these cases. This compares with `r TBI_w %>% filter(((Skull | IC_injury | PTA) & External)) %>% summarise(N=n()) # 26085` identified using the method of Possato et al. The percentage of cases added is at most `r CIs[1] #1.22` with a 95% confidence interval running from `r CIs[2]` to `r CIs[3]`. The specificity of the methods cannot be calculated, as there is no way to manually recheck the coding, however the sensitivity is `r 100 - CIs[1]`.


# Produce the working file

```{r, results = 'show'}
TBI_final<- TBI_w %>%
  filter(((Skull | IC_injury | PTA) & External) | Detail == 'TBI')
  
summary(tableby(External ~ Skull + IC_injury, data=TBI_final, test=FALSE, total=FALSE),text=TRUE)

```

#Save the data files for further work

```{r}
saveRDS(TBI_final, file='data/TBI_final.Rds')

```
