---
title: "Data load"
author: "Anthony Staines & Anne O\\'Farrell"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
library(lubridate)
library(gt)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE,cache = TRUE)
```

# Load files

There are data files for each year from 2013 to October 2020. The rest of 2020 is awaited.

```{r load}
D2013 <- readxl::read_excel('data/2013 head injuries.xls')
D2014 <- readxl::read_excel('data/2014 head injuries.xls')
D2015 <- readxl::read_excel('data/2015 head injuries.xls')
D2016 <- readxl::read_excel('data/2016 head injuries.xls')
D2017 <- readxl::read_excel('data/2017 head injuries.xls')
D2018 <- readxl::read_excel('data/2018 head injuries.xls')
D2019 <- readxl::read_excel('data/2019 head injuries.xls')
D2020 <- readxl::read_excel('data/2020_part head injuries.xls') # Up to October

#names(D2019)
#vtable::vt(D2013)


TBI <- D2020 %>% rbind(D2019) %>% rbind(D2018) %>% rbind(D2017) %>% rbind(D2016) %>% rbind(D2015) %>% rbind(D2014) %>% rbind(D2013)

nrow(TBI)
nrow(distinct(TBI))

TBI <- TBI %>% distinct() #Some dates overlap 2019 and part 2020
rm(D2013,D2014,D2015,D2016,D2017,D2018,D2019,D2020)

```

# Data
 Fix the variable names

```{r FixNames, echo=FALSE}

# Spaces to _
# - to _
# Remove _(Y/N)
# Remove brackets
# / to or

NAMES <- as_tibble(names(TBI))
NAMES %>% write_csv('data/names.csv') # Manually altered
NAMES <- read_csv('data/names_revised.csv')# %>% select(-X1)

names(TBI) <- NAMES$NewName  
```

Fix the variable types

```{r FixTypes}
table(NAMES$NewName,NAMES$Type)

#Numbers
Numbers <- NAMES %>% filter(Type == 'Numeric') %>% select(NewName)
  COLS <- Numbers$NewName
TBIt <- TBI %>%
  mutate(across(all_of(COLS), as.numeric))
  rm(Numbers)

#Dates
Dates <- NAMES%>% filter(Type == 'Date') %>% select(NewName)
  COLS <- Dates$NewName
TBIt <- TBIt %>%
  mutate(across(all_of(COLS), as_date))
  rm(Dates)

#vtable::vt(TBIt)
#warnings()

TBI <- TBIt
rm(TBIt, COLS, NAMES)
```

# Add new variables

```{r AddNew}
#Correctly ordered age groups
TBI <- TBI %>%
  mutate(AgeGrp = as_factor(Age_by_5_year)) %>%
  mutate(AgeGrp = fct_reorder(AgeGrp, Age_by_1_year))#  #order by median age
#  mutate(AgeGrp = fct_relevel(AgeGrp, "5-9", after=1)) # 
#  mutate(AgeGrp = fct_relevel(AgeGrp, "100-104", after=Inf)) #

TBI %>% select(AgeGrp) %>% group_by(AgeGrp) %>% summarise(N = n())

#UniqueID
TBI$ID = 1:nrow(TBI)

TBI <- TBI %>% select(ID,E_MRN:AgeGrp)

```

# Basic tables

```{r}
Gender <- gt(data=TBI %>%
               select(Gender) %>%
               group_by(Gender) %>%
               summarise(N = n())
             ) %>%
  tab_header(
    title = "Gender",
    subtitle = "Only Male and Female are recorded"
  )
Gender

Age <- gt(data=TBI %>%
            select(AgeGrp) %>%
            group_by(AgeGrp) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Age (5 year age groups)",
    subtitle = "All S00-S09 admissions"
  )
Age

Source <- gt(data=TBI %>%
            select(Admission_source) %>%
            group_by(Admission_source) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Source of admission",
    subtitle = "All S00-S09 admissions"
  )
Source

Type <- gt(data=TBI %>%
            select(Admission_type) %>%
            group_by(Admission_type) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Type of admission",
    subtitle = "All S00-S09 admissions"
  )
Type

Destination <- gt(data=TBI %>%
            select(Discharge_destination) %>%
            group_by(Discharge_destination) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Discharge destination",
    subtitle = "All S00-S09 admissions"
  )
Destination

Group.db <- TBI %>%
  select(Medical_card, NTPF, Public_or_private) %>%
  group_by(Medical_card, NTPF, Public_or_private) %>%
  summarise(N=n())

Group.db %>%
  kbl() %>%
  kable_classic(full_width = F, html_font = "Cambria")

rm(Age, Destination, Gender, Group.db, Source, Type)

```

# Basic plots

```{r}
Daily_Admissions <- TBI %>%
  select(Admission_date) %>%
  group_by(Admission_date) %>%
  summarise(N = n()) %>%
  mutate(Day_of_week = wday(Admission_date,
                            label = TRUE)) %>%
  mutate(Weekend = ifelse((Day_of_week %in% c('Sat','Sun')), TRUE, FALSE))
  
str(Daily_Admissions)

ggplot(Daily_Admissions, aes(x=Admission_date, y=N)) +
  geom_point(aes(colour=Weekend)) +
  geom_smooth( method = "gam", formula = y ~ s(x, k=20, bs = "cs"))  +
  geom_line(alpha=0.2,colour='blue') +
  scale_colour_manual(values=c('lightblue','yellow')) +
  scale_x_date(date_minor_breaks='1 month', date_breaks = '3 months', date_labels = '%b-%y', limits = c(as_date('2013-01-01'),NA)  )


rm(Daily_Admissions)
```

Something very different in 2018

Small gap in late 2018 - only given up to December 1st

Sharp dip in March 2020 - COVID1

This is all head injuries.

# Dictionary of codes

This is set of all the ICD10 codes used in the data which were associated with names. Some of these are blank and were filled in form where they were given in the dataset, some completed from the US codes and a few were completed manually.

```{r}
Dictionary <- TBI %>%
  select(ID,
         Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>% # Generate one row per diagnosis
  filter(!is.na(ICD_code)) %>% # Lose blank codes
  distinct(ICD_code,ICD_name) %>%
  mutate(Name_missing = is.na(ICD_name)) %>%
  mutate(Source = ifelse(Name_missing, NA, 'HIPE'))

```

# Working data set


```{r}

# Codes used and the number of the associated diagnosis
Codes_used_by_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Dx,Code,ICD_name) %>% # Group by diagnosis, and code
  summarize(N=n()) %>% # Count every combination of diagnosis and code
  arrange(desc(N)) # Put from most common to least common

Codes_used_by_Dx

# Codes used for all diagnoses
Codes_used_all_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code,ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code used for any diagnosis
  arrange(desc(N)) # Put from most common to least common

Codes_used_all_Dx

#Codes used for diagnosis 1 only
Codes_used_Dx_1 <- TBI %>%
  select(ID,Dx_1_ICD_code) %>%  #Principal diagnosis
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code, ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code for principal diagnosis
  arrange(desc(N)) # Put from most common to least common

Codes_used_Dx_1

rm(Codes_used_all_Dx, Codes_used_by_Dx, Codes_used_Dx_1)
```


# Working file

## Pivot TBI to long -  one row per non-missing diagnosis

Each row is one diagnosis for one HIPE episode. Patients with many diagnoses have many rows.

```{r}
TBI_l <- TBI %>%
  select(ID,E_MRN,Hospital_name,starts_with('Dx')) %>%
  pivot_longer(cols = c(Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name),
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>%
  filter(!is.na(ICD_code))

```

Work through codes in Oz paper
Pozzato I, Tate RL, Rosenkoetter U, Cameron ID. Epidemiology of hospitalised traumatic brain injury in the state of New South Wales, Australia: a population-based study. Australian and New Zealand Journal of Public Health [Internet]. 2019 [cited 2021 Feb 2];43(4):382–8. Available from: https://onlinelibrary.wiley.com/doi/abs/10.1111/1753-6405.12878

## Defining features of injury

### TBI Diagnosis Codes 1

Skull fracture
 Fracture of vault or base of the skull S02.0- S02.1
 Fractures of skull and facial bones of other or unspecified nature S02.7- S02.9

Intracranial injury
 Concussion  S06.00
 Traumatic cerebral oedema S06.1
 Diffuse cerebral/cerebellar contusions or haematomas  S06.20-S06.28
 Focal cerebral/cerebellar contusion or haematoma  S06.30- S06.38
 Epidural/subarachnoid/subdural haemorrhage  S06.4- S06.6
 Intracranial injuries of other or unspecified nature  S06.8- S06.9

Crush injury of the head
 Crushing injury of skull  S07.1
 
```{r}
CODES <- TBI_l %>% 
  select(ICD_code) %>%
  group_by(ICD_code) %>%
  summarise(N = n()) %>%
  arrange(desc(N)) # 6,842 codes used

#
# Skull fracture
# Fracture of vault or base of the skull S02.0- S02.1
# Fractures of skull and facial bones of other or unspecified nature S02.7- S02.9
#
#Intracranial injury
# Concussion  S06.00
# Traumatic cerebral oedema S06.1
# Diffuse cerebral/cerebellar contusions or haematomas  S06.20-S06.28
# Focal cerebral/cerebellar contusion or haematoma  S06.30- S06.38
# Epidural/subarachnoid/subdural haemorrhage  S06.4- S06.6
# Intracranial injuries of other or unspecified nature  S06.8- S06.9
#
#Crush injury of the head
# Crushing injury of skull  S07.1
# 

CODES <- CODES %>%
  mutate(Skull = str_detect(ICD_code, 'S020|S021|S027|S028|S029')) %>%
    mutate(Skull = ifelse(Skull, Skull, NA)) %>%
  mutate(IC_injury = str_detect(ICD_code, 'S060|S061|S062|S063|S064|S065|S066|S068|S069|S071')) %>%
    mutate(IC_injury = ifelse(IC_injury, IC_injury, NA)) %>% 
  mutate(Crush = str_detect(ICD_code, 'S071')) %>%
    mutate(Crush = ifelse(Crush, Crush, NA))
```


### TBI diagnosis Codes #2

Loss of consciousness

 LoC of unspecified duration  S06.01
 LoC < 30 min                 S06.02
 LoC 30 min to 24 hours       S06.03
 LoC > 24 hours               S06.04-05

```{r}
#
#LOC
# LoC of unspecified duration   S06.01
# LoC < 30 min                  S06.02
# LoC 30 min to 24 hours        S06.03
# LoC > 24 hours                S06.04-05
#
CODES <- CODES %>%
  mutate(LOC = str_detect(ICD_code, 'S0601|S0602|S0603|S0604|S0605')) %>%
    mutate(LOC = ifelse(LOC, LOC, NA)) %>%
  mutate(LOC_duration = recode(ICD_code,
                           'S0601' = 'n.s.',
                           'S0602' = '< 0.5',
                           'S0603' = '< 24',
                           'S0604' = '> 24',
                           'S0605' = '> 24',
                           .default = NA_character_
                           ))
```

## Injury-related characteristics

### Other TBI severity codes

#### Post traumatic amnesia

 PTA of unspecified duration   F04.0
 PTA < 24 hours                F04.1
 PTA 24 hours to 2 weeks       F04.2
 PTA ≥ 2 weeks                 F04.3


```{r}
#Post traumatic amnesia
# PTA of unspecified duration   F04.0
# PTA < 24 hours                F04.1
# PTA 24 hours to 2 weeks       F04.2
# PTA ≥ 2 weeks                 F04.3
#
CODES <- CODES %>%
  mutate(PTA = str_detect(ICD_code, 'F0400|F0401|F0402|F0403')) %>%
    mutate(PTA = ifelse(PTA, PTA, NA)) %>%
  mutate(PTA_duration = recode(ICD_code,
                           'F0400' = 'n.s.',
                           'F0401' = '< 1d',
                           'F0402' = '< 2w',
                           'F0403' = '> 2w',
                           .default = NA_character_
                           )) %>%
    select(-N) # Drop the count
```

We've now got a file which classifies each code used in the data set into one of five groups Skull, IC_injury, Crush, LOC, LOC_duration, PTA, and PTA_duration. We join this back into TBI_l"

```{r}
TBI_l <- TBI_l %>%
  left_join(CODES, by = "ICD_code")

rm(CODES)
```

#### E-codes
External cause codes (E-Codes)

Transport accidents   V01-V99
Falls                 W00-W19
Assaults              X85-Y09
Other accidents       W20-W64, X60-X84

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 

```{r}

Dictionary <- Dictionary %>%
  arrange(ICD_code) %>%
  mutate(Row = row_number())

#V01 - V99 inclusive
Transport <- Dictionary %>%
  filter(str_starts(ICD_code, 'V' )) %>%
  mutate(Transport = TRUE) %>%
  select(ICD_code, ICD_name, Transport)

#W00 - W19 Row 5795 to Row 5875
Dictionary %>% filter(str_starts(ICD_code, 'W0'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'W00')) %>%
    select(Row)
  START <- as.numeric(START) #5795
Dictionary %>% filter(str_starts(ICD_code, 'W19'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'W19')) %>%
    select(Row)
  END <- as.numeric(END) # 5875

Falls <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Falls = TRUE)%>%
  select(ICD_code, ICD_name, Falls)

#X85 - Y09
Dictionary %>% filter(str_starts(ICD_code, 'X85'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'X8502')) %>%
    select(Row)
  START <- as.numeric(START) #6065
Dictionary %>% filter(str_starts(ICD_code, 'Y09'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'Y0909')) %>%
    select(Row)
  END <- as.numeric(END) # 6192

Assaults <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Assaults = TRUE) %>%
  select(ICD_code, ICD_name, Assaults)

#W20 - 264 Other accidents 1
Dictionary %>% filter(str_starts(ICD_code, 'W20'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'W20')) %>%
    select(Row)
  START <- as.numeric(START) #5876
Dictionary %>% filter(str_starts(ICD_code, 'W64'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'W64')) %>%
    select(Row)
  END <- as.numeric(END) # 5954

Other1 <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Other = TRUE)

#X60 - X84 Other accidents 2
Dictionary %>% filter(str_starts(ICD_code, 'X60'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'X60')) %>%
    select(Row)
  START <- as.numeric(START) #6025
Dictionary %>% filter(str_starts(ICD_code, 'X84'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'X84')) %>%
    select(Row)
  END <- as.numeric(END) # 6064

Other2 <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Other = TRUE)

Other = Other1 %>%
  rbind(Other2) %>%
  select(ICD_code, ICD_name, Other)

rm(Other1, Other2)

```

# Associated factors
Associated factor codes

 Drug use      F11-F16, F18-F19, X40-4, X60-4, Z72.2
 Alcohol use   F10.0-8, T51.0, T51.9, X45, X65, Y90.1-9, Z72.1   
 Sport         Y92.30-9
 Recreation    U50-U72

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 

```{r}
# Drug use
# F11-F16, F18-F19, X40-4, X60-4, Z72.2
Dictionary %>% filter(str_starts(ICD_code, 'F11'))
Dictionary %>% filter(str_starts(ICD_code, 'F16'))
Dictionary %>% filter(str_starts(ICD_code, 'F18'))
Dictionary %>% filter(str_starts(ICD_code, 'F19'))
                      
Row <- Dictionary %>%
  filter(ICD_code >= 'F110' & ICD_code < 'F170')%>%
  rbind(Dictionary %>%
          filter(str_starts(ICD_code, 'F18') |
                   str_starts(ICD_code, 'F19'))) %>%
  rbind(Dictionary %>%
          filter(ICD_code >= 'X40' & ICD_code < 'X45')) %>%
  rbind(Dictionary %>%
          filter(ICD_code >= 'X60' & ICD_code < 'X65')) %>%
  rbind(Dictionary %>%
          filter(ICD_code == 'Z722'))

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Drug <-  Dictionary %>%
  slice(RowSlice) %>%
  mutate(Drug = TRUE)%>%
  select(ICD_code, ICD_name, Drug)


#Alcohol use
# F10.0-8, T51.0, T51.9, X45, X65, Y90.1-9, Z72.1
Row <- Dictionary %>% filter((str_starts(ICD_code, 'F10') & !str_starts(ICD_code, 'F109')) |
                        str_starts(ICD_code, 'T510') | str_starts(ICD_code, 'T519'))

Row <- Row %>% rbind(
Dictionary %>% filter(str_starts(ICD_code, 'X45') | str_starts(ICD_code, 'X65'))
  )

Row <- Row %>% rbind(
  Dictionary %>% filter(str_starts(ICD_code, 'Y90') | str_starts(ICD_code, 'Z721'))
  )

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Alcohol <-  Dictionary %>%
  slice(RowSlice) %>%
  mutate(Alcohol = TRUE)%>%
  select(ICD_code, ICD_name, Alcohol)

#Sport
#Y9230 - Y9239 Row 6408 to Row 6416
Dictionary %>% filter(str_starts(ICD_code, 'Y923'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'Y9230')) %>%
    select(Row)
  START <- as.numeric(START) #6408
Dictionary %>% filter(str_starts(ICD_code, 'Y923'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'Y9239')) %>%
    select(Row)
  END <- as.numeric(END) # 6416

Sport <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Sport = TRUE)%>%
  select(ICD_code, ICD_name, Sport)

#Recreation
#U50 - U72 Row 5168 to Row 5310
Dictionary %>% filter(str_starts(ICD_code, 'U5'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'U5000')) %>%
    select(Row)
  START <- as.numeric(START) #5168
Dictionary %>% filter(str_starts(ICD_code, 'U7'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'U72')) %>%
    select(Row)
  END <- as.numeric(END) # 5310

Recreation <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Recreation = TRUE)%>%
  select(ICD_code, ICD_name, Recreation)

Dictionary %>% slice(seq(1,5),seq(6,10))
Dictionary %>% slice(seq(1,5),6,seq(7,10))

```

## Associated injury

 Other mechanical trauma   S10-T14
 Complications   T80-88
 Other injuries    T15-T79, T89

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 


```{r}
#Other mechanical trauma
#S10-T14
Dictionary %>% filter(str_starts(ICD_code, 'S10'))
Dictionary %>% filter(str_starts(ICD_code, 'T14')) %>% tail()

                      
Row <- Dictionary %>%
  filter(ICD_code >= 'S100' & ICD_code < 'T15') %>%
  rbind(Dictionary %>%
  filter(ICD_code >= 'T80' & ICD_code < 'T89'))

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Other.mechanical <- Dictionary %>%
  slice(RowSlice) %>%
  mutate(Other.mechanical = TRUE)%>%
  select(ICD_code, ICD_name, Other.mechanical)

#Complications
#T80-88

Dictionary %>% filter(str_starts(ICD_code, 'T8')) 
Dictionary %>% filter(str_starts(ICD_code, 'T88'))

Row <- Dictionary %>%
  filter(ICD_code >= 'T8' & ICD_code < 'T89')

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Complications <- Dictionary %>%
  slice(RowSlice) %>%
  mutate(Complications = TRUE)%>%
  select(ICD_code, ICD_name, Complications)


#Other injuries
#T15-T79, T89
Dictionary %>% filter(str_starts(ICD_code, 'T15'))
Dictionary %>% filter(str_starts(ICD_code, 'T79'))
Dictionary %>% filter(str_starts(ICD_code, 'T89'))


Row <- Dictionary %>%
  filter(ICD_code >= 'T15' & ICD_code < 'T80'| str_starts(ICD_code, 'T89'))

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Other.injuries <- Dictionary %>%
  slice(RowSlice) %>%
  mutate(Other.injuries = TRUE)%>%
  select(ICD_code, ICD_name, Other.injuries)

```


#Merge the codes
Join the external code groups to the individual files
 Transport, Falls, Assaults, Other
 Drug, Alcohol, Sport, Recreation
 Other.mechanical, Complications, Other.injuries

```{r}
names(TBI_l)

TBI_l <- TBI_l %>%
  left_join(Transport,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Falls,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Assaults,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Other,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Drug,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Alcohol,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Sport,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Recreation,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Other.mechanical,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Complications,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Other.injuries,  by = c("ICD_code", "ICD_name"))

names(TBI_l)

rm(Transport, Falls, Assaults, Other,
   Drug, Alcohol, Sport, Recreation,
   Other.mechanical, Complications, Other.injuries,
   Row,END,RowSlice,START)
```


# Widen the file
The main file TBI_l now has one row per diagnosis. For many purposes we will need one row per person.


```{r}
#The subfiles come in numerical order
#

#TBI_0
# This file has the ID, the E_MRN, and all the diagnoses in a row.
#
TBI_0 <- TBI_l %>%
  pivot_wider(id_cols = c(ID, E_MRN), names_from = Dx, values_from = c(ICD_code, ICD_name))

#TBI_1
# Convert the per-diagnosis classifications of ancillary variables to per-person
# Each is a group of up to 30 values, TRUE or NA
# ~as.logical(sum(!is.na(.)))
# The missing values are omitted
# The non-missing values are summed
# The total is converted to a logical value
#
TBI_1 <- TBI_l %>%
  select(ID,E_MRN,Skull:LOC,PTA,Transport:Other.injuries) %>%
  group_by(ID, E_MRN) %>%
  summarise(across(Skull:Other.injuries, ~as.logical(sum(!is.na(.)))))

# Quick check on codes for LOC and PTA duration
unique(TBI_l$LOC_duration)
unique(TBI_l$PTA_duration)

#TBI_2
# There are two duration variables, one for LOC and one for PTA
# Some people have more than one, which is surprising
#
TBI_2 <- TBI_l %>%
  select(ID,E_MRN,LOC_duration,PTA_duration) %>%
  group_by(ID, E_MRN) %>%
  mutate(LOCl = ifelse(is.na(LOC_duration),NA,TRUE)) %>%
  mutate(PTAl = ifelse(is.na(PTA_duration),NA,TRUE))

table(TBI_2$LOC_duration, useNA = 'ifany')
table(TBI_2$PTA_duration, useNA = 'ifany')
table(TBI_2$LOC_duration, TBI_2$PTA_duration, useNA = 'ifany')

#Three have two LOC codes provided
Check <- TBI_2 %>%
  group_by(ID,E_MRN) %>%
  summarise(across(LOCl:PTAl, ~sum(!is.na(.)))) %>%
  filter(LOCl > 1)

Check$ID 
#
# None have two PTA codes
Check <- TBI_2 %>%
  group_by(ID,E_MRN) %>%
  summarise(across(LOCl:PTAl, ~sum(!is.na(.)))) %>%
  filter(PTAl > 1)

Check$ID 
#
#Manual fix
# 31887 CVEATIY Two codes, S0602 and S0601 brief and unspecified, Dx1 and Dx5, Choose brief
# 44727 COE5UI3 Two codes, S0601 and S0602 unspecified and brief, Dx2 and Dx15, Choose unspecified
# 69836 8O#3VF1 Two codes, S0601 and S0602, brief and unspecified, Dx5 and Dx6, Choose brief
#

TBI_2 %>%
  filter(ID == 31887)
TBI_2 %>%
  filter(ID == 44727)
TBI_2 %>%
  filter(ID == 69836)

TBI_2$LOC_duration[TBI_2$ID == 31887 & TBI_2$LOC_duration == 'n.s.'] = NA_character_
TBI_2$LOC_duration[TBI_2$ID == 44727 & TBI_2$LOC_duration == '< 0.5'] = NA_character_
TBI_2$LOC_duration[TBI_2$ID == 69836 & TBI_2$LOC_duration == 'n.s.'] = NA_character_

#Check
# Recalculate logical variables with the new data
#
TBI_2 <- TBI_2 %>%
  group_by(ID, E_MRN) %>%
  mutate(LOCl = ifelse(is.na(LOC_duration),NA,TRUE)) %>%
  mutate(PTAl = ifelse(is.na(PTA_duration),NA,TRUE))

#Three had two LOC codes provided
#Should be fixed by now
#
Check <- TBI_2 %>%
  group_by(ID,E_MRN) %>%
  summarise(across(LOCl:PTAl, ~sum(!is.na(.)))) %>%
  filter(LOCl > 1)

Check$ID 
#Now zero length
#

# Recode the LOC and PTA durations as numbers
#
TBI_2 <- TBI_2 %>%
  mutate(LOCn = recode(LOC_duration,
                       'n.s.' = 0,
                       '< 0.5' = 1,
                       '< 24' = 2,
                       '> 24' = 3,
                       )) %>%
  mutate(PTAn = recode(PTA_duration,
                       'n.s.' = 0,
                       '< 1d' = 1,
                       '< 2w' = 2,
                       '> 2w' = 3,
                       ))

#
#TBI_3
#  Cuts down to files with LOC
#
TBI_3 <- TBI_2 %>%
  select(ID,E_MRN, LOC_duration, LOCl, LOCn) %>%
  filter(!is.na(LOC_duration) | !is.na(LOCl) | !is.na(LOCn))

Check <- TBI_3 %>%
  select(ID,E_MRN) %>%
  group_by(ID,E_MRN) %>%
  summarise(N = n()) %>%
  filter(N != 1)
#All zero

#
#TBI_4
#  Cuts down to files with PTA
#
TBI_4 <- TBI_2 %>%
  select(ID,E_MRN, PTA_duration, PTAl, PTAn) %>%
  filter(!is.na(PTA_duration) | !is.na(PTAl) | !is.na(PTAn))

Check <- TBI_4 %>%
  select(ID,E_MRN) %>%
  group_by(ID,E_MRN) %>%
  summarise(N = n()) %>%
  filter(N != 1)
#All zero

#
#Additional variables from TBI
#
TBI_5 <- TBI %>%
  select(ID,E_MRN,Hospital_name:Team_code_principal,Transfer_from:`Dx_1_name_CCS-IM`,Proc_1_ACHI_name:AgeGrp)


#
#Merge the files
# Diagnoses TBI_0,
# Skull:Other.injuries TB.1,
# LOC TBI_3
# PTA TBI_4
# Admission level variables TBI_5

TBI_w <- TBI_0 %>%
  left_join(TBI_1, by = c("ID", "E_MRN")) %>%
  left_join(TBI_3, by = c("ID", "E_MRN")) %>%
  left_join(TBI_4, by = c("ID", "E_MRN")) %>%
  left_join(TBI_5, by = c("ID", "E_MRN"))
  
  

rm(TBI_0, TBI_1, TBI_2, TBI_3, TBI_4, TBI_5, Check)
```

#Save the data files for further work

```{r}
saveRDS(Dictionary, file='data/Dictionary.Rds')
saveRDS(TBI, file='data/TBI_Rds')
saveRDS(TBI_w, file='data/TBI_w.Rds')
saveRDS(TBI_l, file='data/TBI_l.Rds')

rm(list=ls())
```
