---
title: "Data load"
author: "Anthony Staines & Anne O\\'Farrell"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
library(lubridate)
library(gt)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

# Load files

There are data files for each year from 2013 to October 2020. The rest of 2020 is awaited.

```{r load}
D2013 <- readxl::read_excel('data/2013 head injuries.xls')
D2014 <- readxl::read_excel('data/2014 head injuries.xls')
D2015 <- readxl::read_excel('data/2015 head injuries.xls')
D2016 <- readxl::read_excel('data/2016 head injuries.xls')
D2017 <- readxl::read_excel('data/2017 head injuries.xls')
D2018 <- readxl::read_excel('data/2018 head injuries.xls')
D2019 <- readxl::read_excel('data/2019 head injuries.xls')
D2020 <- readxl::read_excel('data/2020_part head injuries.xls') # Up to October

#names(D2019)
#vtable::vt(D2013)


TBI <- D2020 %>% rbind(D2019) %>% rbind(D2018) %>% rbind(D2017) %>% rbind(D2016) %>% rbind(D2015) %>% rbind(D2014) %>% rbind(D2013)

nrow(TBI)
nrow(distinct(TBI))

TBI <- TBI %>% distinct() #Some dates overlap 2019 and part 2020
rm(D2013,D2014,D2015,D2016,D2017,D2018,D2019,D2020)

```

# Data
 Fix the variable names

```{r FixNames, echo=FALSE}

# Spaces to _
# - to _
# Remove _(Y/N)
# Remove brackets
# / to or

NAMES <- as_tibble(names(TBI))
NAMES %>% write_csv('data/names.csv') # Manually altered
NAMES <- read_csv('data/names_revised.csv')# %>% select(-X1)

names(TBI) <- NAMES$NewName  
```

Fix the variable types

```{r FixTypes}
table(NAMES$NewName,NAMES$Type)

#Numbers
Numbers <- NAMES %>% filter(Type == 'Numeric') %>% select(NewName)
  COLS <- Numbers$NewName
TBIt <- TBI %>%
  mutate(across(all_of(COLS), as.numeric))
  rm(Numbers)

#Dates
Dates <- NAMES%>% filter(Type == 'Date') %>% select(NewName)
  COLS <- Dates$NewName
TBIt <- TBIt %>%
  mutate(across(all_of(COLS), as_date))
  rm(Dates)

#vtable::vt(TBIt)
#warnings()

TBI <- TBIt
rm(TBIt, COLS, NAMES)
```

# Add new variables

```{r AddNew}
#Correctly ordered age groups
TBI <- TBI %>%
  mutate(AgeGrp = as_factor(Age_by_5_year)) %>%
  mutate(AgeGrp = fct_reorder(AgeGrp, Age_by_1_year))#  #order by median age
#  mutate(AgeGrp = fct_relevel(AgeGrp, "5-9", after=1)) # 
#  mutate(AgeGrp = fct_relevel(AgeGrp, "100-104", after=Inf)) #

TBI %>% select(AgeGrp) %>% group_by(AgeGrp) %>% summarise(N = n())

#UniqueID
TBI$ID = 1:nrow(TBI)

TBI <- TBI %>% select(ID,E_MRN:AgeGrp)

```

# Basic tables

```{r}
Gender <- gt(data=TBI %>%
               select(Gender) %>%
               group_by(Gender) %>%
               summarise(N = n())
             ) %>%
  tab_header(
    title = "Gender",
    subtitle = "Only Male and Female are recorded"
  )
Gender

Age <- gt(data=TBI %>%
            select(AgeGrp) %>%
            group_by(AgeGrp) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Age (5 year age groups)",
    subtitle = "All S00-S09 admissions"
  )
Age

Source <- gt(data=TBI %>%
            select(Admission_source) %>%
            group_by(Admission_source) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Source of admission",
    subtitle = "All S00-S09 admissions"
  )
Source

Type <- gt(data=TBI %>%
            select(Admission_type) %>%
            group_by(Admission_type) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Type of admission",
    subtitle = "All S00-S09 admissions"
  )
Type

Destination <- gt(data=TBI %>%
            select(Discharge_destination) %>%
            group_by(Discharge_destination) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Discharge destination",
    subtitle = "All S00-S09 admissions"
  )
Destination

Group.db <- TBI %>%
  select(Medical_card, NTPF, Public_or_private) %>%
  group_by(Medical_card, NTPF, Public_or_private) %>%
  summarise(N=n())

Group.db %>%
  kbl() %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

# Basic plots

```{r}
Daily_Admissions <- TBI %>%
  select(Admission_date) %>%
  group_by(Admission_date) %>%
  summarise(N = n()) %>%
  mutate(Day_of_week = wday(Admission_date,
                            label = TRUE)) %>%
  mutate(Weekend = ifelse((Day_of_week %in% c('Sat','Sun')), TRUE, FALSE))
  
str(Daily_Admissions)

ggplot(Daily_Admissions, aes(x=Admission_date, y=N)) +
  geom_point(aes(colour=Weekend)) +
  geom_smooth( method = "gam", formula = y ~ s(x, k=20, bs = "cs"))  +
  geom_line(alpha=0.2,colour='blue') +
  scale_colour_manual(values=c('lightblue','yellow')) +
  scale_x_date(date_minor_breaks='1 month', date_breaks = '3 months', date_labels = '%b-%y', limits = c(as_date('2013-01-01'),NA)  )

```

Something very different in 2018

Small gap in late 2018 - only given up to December 1st

Sharp dip in March 2020 - COVID1

This is all head injuries.

# Dictionary of codes

This is set of all the ICD10 codes used in the data which were associated with names. Some of these are blank and were filled in form where they were given in the dataset, some completed from the US codes and a few were completed manually.

```{r}
Dictionary <- TBI %>%
  select(ID,
         Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>% # Generate one row per diagnosis
  filter(!is.na(ICD_code)) %>% # Lose blank codes
  distinct(ICD_code,ICD_name) %>%
  mutate(Name_missing = is.na(ICD_name)) %>%
  mutate(Source = ifelse(Name_missing, NA, 'HIPE'))

```

# Working data set

Advice?
Suggest, exclude all electives.
Require SO6 as Dx_1 or SO6 + Vcode

```{r}

# Codes used and the number of the associated diagnosis
Codes_used_by_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Dx,Code,ICD_name) %>% # Group by diagnosis, and code
  summarize(N=n()) %>% # Count every combination of diagnosis and code
  arrange(desc(N)) # Put from most common to least common

Codes_used_by_Dx

# Codes used for all diagnoses
Codes_used_all_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code,ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code used for any diagnosis
  arrange(desc(N)) # Put from most common to least common

Codes_used_all_Dx

#Codes used for diagnosis 1 only
Codes_used_Dx_1 <- TBI %>%
  select(ID,Dx_1_ICD_code) %>%  #Principal diagnosis
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code, ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code for principal diagnosis
  arrange(desc(N)) # Put from most common to least common

Codes_used_Dx_1
```


# Working file

## Pivot TBI to long -  one row per non-missing diagnosis

Each row is one diagnosis for one HIPE episode. Patients with many diagnoses have many rows.

```{r}
TBIl <- TBI %>%
  select(ID,E_MRN,Hospital_name,starts_with('Dx')) %>%
  pivot_longer(cols = c(Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name),
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>%
  filter(!is.na(ICD_code))

```

Work though codes in Oz paper
Pozzato I, Tate RL, Rosenkoetter U, Cameron ID. Epidemiology of hospitalised traumatic brain injury in the state of New South Wales, Australia: a population-based study. Australian and New Zealand Journal of Public Health [Internet]. 2019 [cited 2021 Feb 2];43(4):382–8. Available from: https://onlinelibrary.wiley.com/doi/abs/10.1111/1753-6405.12878


```{r}
CODES <- TBIl %>% 
  select(ICD_code) %>%
  group_by(ICD_code) %>%
  summarise(N = n()) %>%
  arrange(desc(N)) # 6,842 codes used

CODES <- CODES %>%
  mutate(Skull = str_detect(ICD_code, 'S020|S021|S027|S028|S029')) %>%
  mutate(IC_Injury = str_detect(ICD_code, 'S060|S061|S062|S063|S064|S065|S066|S068|S069|S071')) %>%
  mutate(LOC = str_detect(ICD_code, 'S0601|S0602|S0603|S0604|S0605')) %>%
  mutate(LOC.Duration = recode(ICD_code,
                           'S0601' = 'n.s. ',
                           'S0602' = '< 0.5',
                           'S0603' = '< 24 ',
                           'S0604' = '> 24 ',
                           'S0605' = '> 24 ',
                           .default = NA_character_
                           )) %>%
  mutate(PTA = str_detect(ICD_code, 'F0400|F0401|F0402|F0403')) %>%
  mutate(PTA.Duration = recode(ICD_code,
                           'F0400' = 'n.s. ',
                           'F0401' = '< 1d',
                           'F0402' = '< 2w ',
                           'F0403' = '> 2w ',
                           .default = NA_character_
                           )) 
TBIl2 <- TBIl %>%
  left_join(CODES, by = "ICD_code")
```

Working file

```{r}
TBIw <- TBI %>%
  # Lose electives, maternity and newborns
  filter(Admission_type == 'Emergency' |
           Admission_type == 'Emergency Readmission') 
  
TBIl <- TBIl %>%
  mutate(S06 = str_detect(string = ICD_code, pattern = '^S06')) %>%
  mutate(IC_Injury = str_detect(string = `Dx_1_name_CCS-IM`,
                                pattern = 'Intracranial injury')) %>%
  mutate(ELIGIBLE = (S06 | IC_Injury))

TBIw <- TBIw %>%
  left_join(TBIl %>% filter(ELIGIBLE) %>% select(ID,ELIGIBLE) %>% distinct(), by = "ID")

```
