---
title: "Data load"
author: "Anthony Staines & Anne O\\'Farrell"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
library(lubridate)
library(gt)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

# Load files

There are data files for each year from 2013 to October 2020. The rest of 2020 is awaited.

```{r load}
D2013 <- readxl::read_excel('data/2013 head injuries.xls')
D2014 <- readxl::read_excel('data/2014 head injuries.xls')
D2015 <- readxl::read_excel('data/2015 head injuries.xls')
D2016 <- readxl::read_excel('data/2016 head injuries.xls')
D2017 <- readxl::read_excel('data/2017 head injuries.xls')
D2018 <- readxl::read_excel('data/2018 head injuries.xls')
D2019 <- readxl::read_excel('data/2019 head injuries.xls')
D2020 <- readxl::read_excel('data/2020_part head injuries.xls') # Up to October

#names(D2019)
#vtable::vt(D2013)


TBI <- D2020 %>% rbind(D2019) %>% rbind(D2018) %>% rbind(D2017) %>% rbind(D2016) %>% rbind(D2015) %>% rbind(D2014) %>% rbind(D2013)

nrow(TBI)
nrow(distinct(TBI))

TBI <- TBI %>% distinct() #Some dates overlap 2019 and part 2020
rm(D2013,D2014,D2015,D2016,D2017,D2018,D2019,D2020)

```

# Data
 Fix the variable names

```{r FixNames, echo=FALSE}

# Spaces to _
# - to _
# Remove _(Y/N)
# Remove brackets
# / to or

NAMES <- as_tibble(names(TBI))
NAMES %>% write_csv('data/names.csv') # Manually altered
NAMES <- read_csv('data/names_revised.csv')# %>% select(-X1)

names(TBI) <- NAMES$NewName  
```

Fix the variable types

```{r FixTypes}
table(NAMES$NewName,NAMES$Type)

#Numbers
Numbers <- NAMES %>% filter(Type == 'Numeric') %>% select(NewName)
  COLS <- Numbers$NewName
TBIt <- TBI %>%
  mutate(across(all_of(COLS), as.numeric))
  rm(Numbers)

#Dates
Dates <- NAMES%>% filter(Type == 'Date') %>% select(NewName)
  COLS <- Dates$NewName
TBIt <- TBIt %>%
  mutate(across(all_of(COLS), as_date))
  rm(Dates)

#vtable::vt(TBIt)
#warnings()

TBI <- TBIt
rm(TBIt, COLS, NAMES)
```

# Add new variables

```{r AddNew}
#Correctly ordered age groups
TBI <- TBI %>%
  mutate(AgeGrp = as_factor(Age_by_5_year)) %>%
  mutate(AgeGrp = fct_reorder(AgeGrp, Age_by_1_year))#  #order by median age
#  mutate(AgeGrp = fct_relevel(AgeGrp, "5-9", after=1)) # 
#  mutate(AgeGrp = fct_relevel(AgeGrp, "100-104", after=Inf)) #

TBI %>% select(AgeGrp) %>% group_by(AgeGrp) %>% summarise(N = n())

#UniqueID
TBI$ID = 1:nrow(TBI)

TBI <- TBI %>% select(ID,E_MRN:AgeGrp)

```

# Basic tables

```{r}
Gender <- gt(data=TBI %>%
               select(Gender) %>%
               group_by(Gender) %>%
               summarise(N = n())
             ) %>%
  tab_header(
    title = "Gender",
    subtitle = "Only Male and Female are recorded"
  )
Gender

Age <- gt(data=TBI %>%
            select(AgeGrp) %>%
            group_by(AgeGrp) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Age (5 year age groups)",
    subtitle = "All S00-S09 admissions"
  )
Age

Source <- gt(data=TBI %>%
            select(Admission_source) %>%
            group_by(Admission_source) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Source of admission",
    subtitle = "All S00-S09 admissions"
  )
Source

Type <- gt(data=TBI %>%
            select(Admission_type) %>%
            group_by(Admission_type) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Type of admission",
    subtitle = "All S00-S09 admissions"
  )
Type

Destination <- gt(data=TBI %>%
            select(Discharge_destination) %>%
            group_by(Discharge_destination) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Discharge destination",
    subtitle = "All S00-S09 admissions"
  )
Destination

Group.db <- TBI %>%
  select(Medical_card, NTPF, Public_or_private) %>%
  group_by(Medical_card, NTPF, Public_or_private) %>%
  summarise(N=n())

Group.db %>%
  kbl() %>%
  kable_classic(full_width = F, html_font = "Cambria")

rm(Age, Destination, Gender, Group.db, Source, Type)

```

# Basic plots

```{r}
Daily_Admissions <- TBI %>%
  select(Admission_date) %>%
  group_by(Admission_date) %>%
  summarise(N = n()) %>%
  mutate(Day_of_week = wday(Admission_date,
                            label = TRUE)) %>%
  mutate(Weekend = ifelse((Day_of_week %in% c('Sat','Sun')), TRUE, FALSE))
  
str(Daily_Admissions)

ggplot(Daily_Admissions, aes(x=Admission_date, y=N)) +
  geom_point(aes(colour=Weekend)) +
  geom_smooth( method = "gam", formula = y ~ s(x, k=20, bs = "cs"))  +
  geom_line(alpha=0.2,colour='blue') +
  scale_colour_manual(values=c('lightblue','yellow')) +
  scale_x_date(date_minor_breaks='1 month', date_breaks = '3 months', date_labels = '%b-%y', limits = c(as_date('2013-01-01'),NA)  )


rm(Daily_Admissions)
```

Something very different in 2018

Small gap in late 2018 - only given up to December 1st

Sharp dip in March 2020 - COVID1

This is all head injuries.

# Dictionary of codes

This is set of all the ICD10 codes used in the data which were associated with names. Some of these are blank and were filled in form where they were given in the dataset, some completed from the US codes and a few were completed manually.

```{r}
Dictionary <- TBI %>%
  select(ID,
         Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>% # Generate one row per diagnosis
  filter(!is.na(ICD_code)) %>% # Lose blank codes
  distinct(ICD_code,ICD_name) %>%
  mutate(Name_missing = is.na(ICD_name)) %>%
  mutate(Source = ifelse(Name_missing, NA, 'HIPE'))

```

# Working data set

Advice?
Suggest, exclude all electives.
Require SO6 as Dx_1 or SO6 + Vcode

```{r}

# Codes used and the number of the associated diagnosis
Codes_used_by_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Dx,Code,ICD_name) %>% # Group by diagnosis, and code
  summarize(N=n()) %>% # Count every combination of diagnosis and code
  arrange(desc(N)) # Put from most common to least common

Codes_used_by_Dx

# Codes used for all diagnoses
Codes_used_all_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code,ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code used for any diagnosis
  arrange(desc(N)) # Put from most common to least common

Codes_used_all_Dx

#Codes used for diagnosis 1 only
Codes_used_Dx_1 <- TBI %>%
  select(ID,Dx_1_ICD_code) %>%  #Principal diagnosis
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code, ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code for principal diagnosis
  arrange(desc(N)) # Put from most common to least common

Codes_used_Dx_1

rm(Codes_used_all_Dx, Codes_used_by_Dx, Codes_used_Dx_1)
```


# Working file

## Pivot TBI to long -  one row per non-missing diagnosis

Each row is one diagnosis for one HIPE episode. Patients with many diagnoses have many rows.

```{r}
TBIl <- TBI %>%
  select(ID,E_MRN,Hospital_name,starts_with('Dx')) %>%
  pivot_longer(cols = c(Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name),
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>%
  filter(!is.na(ICD_code))

```

Work though codes in Oz paper
Pozzato I, Tate RL, Rosenkoetter U, Cameron ID. Epidemiology of hospitalised traumatic brain injury in the state of New South Wales, Australia: a population-based study. Australian and New Zealand Journal of Public Health [Internet]. 2019 [cited 2021 Feb 2];43(4):382–8. Available from: https://onlinelibrary.wiley.com/doi/abs/10.1111/1753-6405.12878


```{r}
CODES <- TBIl %>% 
  select(ICD_code) %>%
  group_by(ICD_code) %>%
  summarise(N = n()) %>%
  arrange(desc(N)) # 6,842 codes used

CODES <- CODES %>%
  mutate(Skull = str_detect(ICD_code, 'S020|S021|S027|S028|S029')) %>%
  mutate(IC_Injury = str_detect(ICD_code, 'S060|S061|S062|S063|S064|S065|S066|S068|S069|S071')) %>%
  mutate(LOC = str_detect(ICD_code, 'S0601|S0602|S0603|S0604|S0605')) %>%
  mutate(LOC.Duration = recode(ICD_code,
                           'S0601' = 'n.s. ',
                           'S0602' = '< 0.5',
                           'S0603' = '< 24 ',
                           'S0604' = '> 24 ',
                           'S0605' = '> 24 ',
                           .default = NA_character_
                           )) %>%
  mutate(PTA = str_detect(ICD_code, 'F0400|F0401|F0402|F0403')) %>%
  mutate(PTA.Duration = recode(ICD_code,
                           'F0400' = 'n.s. ',
                           'F0401' = '< 1d',
                           'F0402' = '< 2w ',
                           'F0403' = '> 2w ',
                           .default = NA_character_
                           )) %>%
  select(-N) # Drop the count

TBIl <- TBIl %>%
  left_join(CODES, by = "ICD_code")

rm(Codes)
```

E-codes
External cause codes (E-Codes)

Transport accidents   V01-V99
Falls                 W00-W19
Assaults              X85-Y09
Other accidents       W20-W64, X60-X84

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 

```{r}

Dictionary <- Dictionary %>%
  arrange(ICD_code) %>%
  mutate(Row = row_number())

#V01 - V99 inclusive
Transport <- Dictionary %>%
  filter(str_starts(ICD_code, 'V' )) %>%
  mutate(Transport = TRUE) %>%
  select(ICD_code, ICD_name, Transport)

#W00 - W19 Row 5795 to Row 5875
Dictionary %>% filter(str_starts(ICD_code, 'W0'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'W00')) %>%
    select(Row)
  START <- as.numeric(START) #5795
Dictionary %>% filter(str_starts(ICD_code, 'W19'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'W19')) %>%
    select(Row)
  END <- as.numeric(END) # 5875

Falls <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Falls = TRUE)%>%
  select(ICD_code, ICD_name, Falls)

#X85 - Y09
Dictionary %>% filter(str_starts(ICD_code, 'X85'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'X8502')) %>%
    select(Row)
  START <- as.numeric(START) #6065
Dictionary %>% filter(str_starts(ICD_code, 'Y09'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'Y0909')) %>%
    select(Row)
  END <- as.numeric(END) # 6192

Assaults <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Assaults = TRUE) %>%
  select(ICD_code, ICD_name, Assaults)

#W20 - 264 Other accidents 1
Dictionary %>% filter(str_starts(ICD_code, 'W20'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'W20')) %>%
    select(Row)
  START <- as.numeric(START) #5876
Dictionary %>% filter(str_starts(ICD_code, 'W64'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'W64')) %>%
    select(Row)
  END <- as.numeric(END) # 5954

Other1 <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Other = TRUE)

#X60 - X84 Other accidents 2
Dictionary %>% filter(str_starts(ICD_code, 'X60'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'X60')) %>%
    select(Row)
  START <- as.numeric(START) #6025
Dictionary %>% filter(str_starts(ICD_code, 'X84'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'X84')) %>%
    select(Row)
  END <- as.numeric(END) # 6064

Other2 <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Other = TRUE)

Other = Other1 %>%
  rbind(Other2) %>%
  select(ICD_code, ICD_name, Other)

rm(Other1, Other2)

```

# Associated factors
Associated factor codes

Drug use      F11-F16, F18-F19, X40-4, X60-4, Z72.2
Alcohol use   F10.0-8, T51.0, T51.9, X45, X65, Y90.1-9, Z72.1   
Sport         Y92.30-9
Recreation    U50-U72

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 

```{r}
# Drug use
# F11-F16, F18-F19, X40-4, X60-4, Z72.2
Dictionary %>% filter(str_starts(ICD_code, 'F11'))
Dictionary %>% filter(str_starts(ICD_code, 'F16'))
Dictionary %>% filter(str_starts(ICD_code, 'F18'))
Dictionary %>% filter(str_starts(ICD_code, 'F19'))
                      
Row <- Dictionary %>%
  filter(ICD_code >= 'F110' & ICD_code < 'F170')%>%
  rbind(Dictionary %>%
          filter(str_starts(ICD_code, 'F18') |
                   str_starts(ICD_code, 'F19'))) %>%
  rbind(Dictionary %>%
          filter(ICD_code >= 'X40' & ICD_code < 'X45')) %>%
  rbind(Dictionary %>%
          filter(ICD_code >= 'X60' & ICD_code < 'X65')) %>%
  rbind(Dictionary %>%
          filter(ICD_code == 'Z722'))

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Drug <-  Dictionary %>%
  slice(RowSlice) %>%
  mutate(Drug = TRUE)%>%
  select(ICD_code, ICD_name, Drug)


#Alcohol use
# F10.0-8, T51.0, T51.9, X45, X65, Y90.1-9, Z72.1
Row <- Dictionary %>% filter((str_starts(ICD_code, 'F10') & !str_starts(ICD_code, 'F109')) |
                        str_starts(ICD_code, 'T510') | str_starts(ICD_code, 'T519'))

Row <- Row %>% rbind(
Dictionary %>% filter(str_starts(ICD_code, 'X45') | str_starts(ICD_code, 'X65'))
  )

Row <- Row %>% rbind(
  Dictionary %>% filter(str_starts(ICD_code, 'Y90') | str_starts(ICD_code, 'Z721'))
  )

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Alcohol <-  Dictionary %>%
  slice(RowSlice) %>%
  mutate(Alcohol = TRUE)%>%
  select(ICD_code, ICD_name, Alcohol)

#Sport
#Y9230 - Y9239 Row 6408 to Row 6416
Dictionary %>% filter(str_starts(ICD_code, 'Y923'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'Y9230')) %>%
    select(Row)
  START <- as.numeric(START) #6408
Dictionary %>% filter(str_starts(ICD_code, 'Y923'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'Y9239')) %>%
    select(Row)
  END <- as.numeric(END) # 6416

Sport <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Sport = TRUE)%>%
  select(ICD_code, ICD_name, Sport)

#Recreation
#U50 - U72 Row 5168 to Row 5310
Dictionary %>% filter(str_starts(ICD_code, 'U5'))
  START <- Dictionary %>%
    filter(str_starts(ICD_code, 'U5000')) %>%
    select(Row)
  START <- as.numeric(START) #5168
Dictionary %>% filter(str_starts(ICD_code, 'U7'))
  END <- Dictionary %>%
    filter(str_starts(ICD_code, 'U72')) %>%
    select(Row)
  END <- as.numeric(END) # 5310

Recreation <- Dictionary %>%
  slice(seq(START, END, by =1)) %>%
  mutate(Recreation = TRUE)%>%
  select(ICD_code, ICD_name, Recreation)

Dictionary %>% slice(seq(1,5),seq(6,10))
Dictionary %>% slice(seq(1,5),6,seq(7,10))

```

#Associated injury
Other mechanical trauma   S10-T14
Complications   T80-88
Other injuries    T15-T79, T89

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 


```{r}
#Other mechanical trauma
#S10-T14
Dictionary %>% filter(str_starts(ICD_code, 'S10'))
Dictionary %>% filter(str_starts(ICD_code, 'T14')) %>% tail()

                      
Row <- Dictionary %>%
  filter(ICD_code >= 'S100' & ICD_code < 'T15') %>%
  rbind(Dictionary %>%
  filter(ICD_code >= 'T80' & ICD_code < 'T89'))

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Other.mechanical <- Dictionary %>%
  slice(RowSlice) %>%
  mutate(Other.mechanical = TRUE)%>%
  select(ICD_code, ICD_name, Other.mechanical)

#Complications
#T80-88

Dictionary %>% filter(str_starts(ICD_code, 'T8')) 
Dictionary %>% filter(str_starts(ICD_code, 'T88'))

Row <- Dictionary %>%
  filter(ICD_code >= 'T8' & ICD_code < 'T89')

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Complications <- Dictionary %>%
  slice(RowSlice) %>%
  mutate(Complications = TRUE)%>%
  select(ICD_code, ICD_name, Complications)


#Other injuries
#T15-T79, T89
Dictionary %>% filter(str_starts(ICD_code, 'T15'))
Dictionary %>% filter(str_starts(ICD_code, 'T79'))
Dictionary %>% filter(str_starts(ICD_code, 'T89'))


Row <- Dictionary %>%
  filter(ICD_code >= 'T15' & ICD_code < 'T80'| str_starts(ICD_code, 'T89'))

RowSlice <- Row %>% select(Row)
  RowSlice <- RowSlice$Row

Other.injuries <- Dictionary %>%
  slice(RowSlice) %>%
  mutate(Other.injuries = TRUE)%>%
  select(ICD_code, ICD_name, Other.injuries)

```


Join the external code groups to the individual files
Transport, Falls, Assaults, Other
Drug, Alcohol, Sport, Recreation
Other.mechanical, Complications, Other.injuries

```{r}
names(TBIl)

TBIl2 <- TBIl %>%
  left_join(Transport,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Falls,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Assaults,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Other,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Drug,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Alcohol,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Sport,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Recreation,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Other.mechanical,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Complications,  by = c("ICD_code", "ICD_name")) %>%
  left_join(Other.injuries,  by = c("ICD_code", "ICD_name"))

names(TBIl2)

rm(Row,CODES,Transport, Falls, Assaults, Other1, Other2, Other,
   Transport, Falls, Assaults, Other,
   Drug, Alcohol, Sport, Recreation,
   Other.mechanical, Complications, Other.injuries,
   END,RowSlice,START)
```


#Working file

```{r}
TBIw <- TBI %>%
  # Lose electives, maternity and newborns
  filter(Admission_type == 'Emergency' |
           Admission_type == 'Emergency Readmission') 
  
TBIl <- TBIl %>%
  mutate(S06 = str_detect(string = ICD_code, pattern = '^S06')) %>%
  mutate(IC_Injury = str_detect(string = `Dx_1_name_CCS-IM`,
                                pattern = 'Intracranial injury')) %>%
  mutate(ELIGIBLE = (S06 | IC_Injury))

TBIw <- TBIw %>%
  left_join(TBIl %>% filter(ELIGIBLE) %>% select(ID,ELIGIBLE) %>% distinct(), by = "ID")

```
