---
title: "Data load"
author: "Anthony Staines & Anne O\\'Farrell"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
library(lubridate)
library(gt)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE,cache = TRUE, message= FALSE, results = 'hide')
```

# Load files

There are data files for each year from 2013 to October 2020. The rest of 2020 is awaited.

```{r load}
D2013 <- readxl::read_excel('data/2013 head injuries.xls') %>% mutate(SOURCE = 'D2013')
D2014 <- readxl::read_excel('data/2014 head injuries.xls') %>% mutate(SOURCE = 'D2014')
D2015 <- readxl::read_excel('data/2015 head injuries.xls') %>% mutate(SOURCE = 'D2015')
D2016 <- readxl::read_excel('data/2016 head injuries.xls') %>% mutate(SOURCE = 'D2016')
D2017 <- readxl::read_excel('data/2017 head injuries.xls') %>% mutate(SOURCE = 'D2017')
D2018_original <- readxl::read_excel('data/2018 head injuries.xls') %>% mutate(SOURCE = 'D2018_original')
D2019 <- readxl::read_excel('data/2019 head injuries.xls') %>% mutate(SOURCE = 'D2019')
D2020 <- readxl::read_excel('data/2020_part head injuries.xls') %>% mutate(SOURCE = 'D2020') # Up to October

D2018_revised  <- readxl::read_excel('data/2018 Head Injuries_revised.xls') %>%
  anti_join(D2018_original) %>% mutate(SOURCE = 'D2018_revised') # Extra 5073 entries

names(D2019)
#vtable::vt(D2013)

TBI <- D2020 %>% rbind(D2019) %>% rbind(D2018_original) %>% rbind(D2017) %>% rbind(D2016) %>% rbind(D2015) %>% rbind(D2014) %>% rbind(D2013)

nrow(TBI) #95422
nrow(distinct(TBI)) # SAME
nrow(distinct(TBI %>% select(-SOURCE))) # 93346

nrow(distinct(TBI)) -nrow(distinct(TBI %>% select(-SOURCE))) # 2096 from 2019 October - December.

TBI <- TBI %>% rbind(D2018_revised) # Add 5073 missed from 2018 before.

nrow(TBI) #100515
nrow(distinct(TBI)) # SAME
nrow(distinct(TBI %>% select(-SOURCE))) # 98419

nrow(distinct(TBI)) - nrow(distinct(TBI %>% select(-SOURCE))) # 2096 from 2019 October - December No extras from 2018

TBI <- TBI %>% distinct(across(1:173), .keep_all = TRUE) #Some dates overlap 2019 and part 2020
names(TBI) # Last one is SOURCE
nrow(TBI) # 98419

rm(D2013,D2014,D2015,D2016,D2017,D2019,D2020)
rm(D2018_original, D2018_revised)
```

# Data
 Fix the variable names

```{r FixNames, echo=FALSE}

# Spaces to _
# - to _
# Remove _(Y/N)
# Remove brackets
# / to or

NAMES <- as_tibble(names(TBI))
NAMES %>% write_csv('data/names.csv') # Manually altered
NAMES <- read_csv('data/names_revised.csv')# %>% select(-X1)

names(TBI) <- NAMES$NewName # 174 names
glimpse(TBI)
```

Fix the variable types

```{r FixTypes}
table(NAMES$NewName,NAMES$Type)

#Numbers
Numbers <- NAMES %>% filter(Type == 'Numeric') %>% select(NewName)
  COLS <- Numbers$NewName
TBIt <- TBI %>%
  mutate(across(all_of(COLS), as.numeric))
  rm(Numbers)

#Dates
Dates <- NAMES %>% filter(Type == 'Date') %>% select(NewName)
  COLS <- Dates$NewName
TBIt <- TBIt %>%
  mutate(across(all_of(COLS), as_date))
  rm(Dates)

#vtable::vt(TBIt)
#warnings()

TBI <- TBIt
rm(TBIt, COLS, NAMES)
```

# Add new variables

```{r AddNew}
#UniqueID
TBI$ID = 1:nrow(TBI)

#Correctly ordered age groups
TBI <- TBI %>%
  mutate(Age_group = as_factor(Age_by_5_year)) %>%
  mutate(Age_group = fct_reorder(Age_group, Age_by_1_year)) %>%
  mutate(Age_group = fct_other(Age_group,
                               drop = c('95-99','100-104','110-114'),
                               other_level='95+')) # Merge top 3 levels into 1

table(TBI$Age_group)
#  #order by median age
#  mutate(Age_group = fct_relevel(Age_group, "5-9", after=1)) # 
#  mutate(Age_group = fct_relevel(Age_group, "100-104", after=Inf)) #

TBI %>% select(Age_group) %>% group_by(Age_group) %>% summarise(N = n())
```

```{r CHECKID, results = 'show'}
names(TBI)
table(TBI$SOURCE)
head(TBI$ID)
#Lots of Repeated MRNs
TBI %>% group_by(E_MRN) %>% summarise(Count = n()) %>% filter(Count > 1) %>% summarise(N=n())
TBI %>% group_by(SOURCE,E_MRN) %>% summarise(Count = n()) %>% filter(Count > 1) %>% summarise(N=n())
TBI %>% group_by(ID) %>% summarise(Count = n()) %>% filter(Count > 1) %>% summarise(N=n())

```

```{r REORDER}
TBI <- TBI %>% dplyr::select(ID,SOURCE,E_MRN:Readm_info,Age_group)


```

# Basic tables

```{r, results = 'show'}
Gender <- gt(data=TBI %>%
               select(Gender) %>%
               group_by(Gender) %>%
               summarise(N = n())
             ) %>%
  tab_header(
    title = "Gender",
    subtitle = "Only Male and Female are recorded"
  )
Gender

Age <- gt(data=TBI %>%
            select(Age_group) %>%
            group_by(Age_group) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Age (5 year age groups)",
    subtitle = "All S00-S09 admissions"
  )
Age

Source <- gt(data=TBI %>%
            select(Admission_source) %>%
            group_by(Admission_source) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Source of admission",
    subtitle = "All S00-S09 admissions"
  )
Source

Type <- gt(data=TBI %>%
            select(Admission_type) %>%
            group_by(Admission_type) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Type of admission",
    subtitle = "All S00-S09 admissions"
  )
Type

Destination <- gt(data=TBI %>%
            select(Discharge_destination) %>%
            group_by(Discharge_destination) %>%
            summarise(N = n())
             ) %>%
  tab_header(
    title = "Discharge destination",
    subtitle = "All S00-S09 admissions"
  )
Destination

Group.db <- TBI %>%
  select(Medical_card, NTPF, Public_or_private) %>%
  group_by(Medical_card, NTPF, Public_or_private) %>%
  summarise(N=n())

Group.db %>%
  kbl() %>%
  kable_classic(full_width = F, html_font = "Cambria")

rm(Age, Destination, Gender, Group.db, Source, Type)

```

# Basic plots

```{r}
Daily_Admissions <- TBI %>%
  select(Admission_date) %>%
  group_by(Admission_date) %>%
  summarise(N = n()) %>%
  mutate(Day_of_week = wday(Admission_date,
                            label = TRUE)) %>%
  mutate(Weekend = ifelse((Day_of_week %in% c('Sat','Sun')), TRUE, FALSE))
  
str(Daily_Admissions)

ggplot(Daily_Admissions, aes(x=Admission_date, y=N)) +
  geom_point(aes(colour=Weekend)) +
  geom_smooth( method = "gam", formula = y ~ s(x, k=20, bs = "cs"))  +
  geom_line(alpha=0.2,colour='blue') +
  scale_colour_manual(values=c('lightblue','yellow')) +
  scale_x_date(date_minor_breaks='1 month',
               date_breaks = '3 months',
               date_labels = '%b-%y') +
  labs(title = 'Count of daily admissions for ICD-10 codes S00 - S99',
       subtitle = 'GAM smooth added',
       x = 'Admission date',
       y = 'Cases per day',
       caption='Source HIPE data') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = -90))


rm(Daily_Admissions)
```

This is all head injuries with codes in S00 to S99.

# Dictionary of codes

This is set of all the ICD10 codes used in the data which were associated with names. Some of these are blank and were filled in from where they were given in the dataset, some completed from the US codes and a few were completed manually.

```{r}
Dictionary <- TBI %>%
  select(ID,
         Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>% # Generate one row per diagnosis
  filter(!is.na(ICD_code)) %>% # Lose blank codes
  distinct(ICD_code,ICD_name) %>%
  mutate(Name_missing = is.na(ICD_name)) %>%
  mutate(Data_source = ifelse(Name_missing, NA, 'HIPE'))

```

# ICD-10 Codes used


## Codes used for diagnosis 1 only

```{r, results='asis'}
Codes_used_Dx_1 <- TBI %>%
  select(ID,Dx_1_ICD_code) %>%  #Principal diagnosis
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code, ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code for principal diagnosis
  arrange(desc(N)) # Put from most common to least common

pander::pander(head(Codes_used_Dx_1,n=10))
```

The commonest code used in the first diagnosis was 'Unspecified injury of head', with  'Open wound of other parts of the head', 'Fracture of nasal bones' and 'Other specified injuries of head' roughly equally common.

## Codes used for all diagnoses

```{r, results='asis'}
Codes_used_all_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Code,ICD_name) %>% # Group by code
  summarize(N=n()) %>% # Count every code used for any diagnosis
  arrange(desc(N)) # Put from most common to least common

pander::pander(head(Codes_used_all_Dx,n=10))
```

Across all diagnoses, the commonest codes used were 'Unspecified activity', 'Unspecified place of occurrence' and 'Unspecified fall', suggesting that in many medical records these data were not given.

## Codes used and the number of the associated diagnosis

```{r, results='asis'}
Codes_used_by_Dx <- TBI %>%
  select(ID,Dx_1_ICD_code:Dx_30_ICD_code) %>%  #All diagnoses
  pivot_longer(-ID,
               names_to=c("Dx","Type"),
               names_pattern="(Dx_\\d+)_(.*)",
               values_to="Code") %>% # Generate one row per diagnosis
  select(-Type) %>% # This just says ICD_code
  filter(!is.na(Code)) %>% # Lose blank codes
  left_join(Dictionary, by = c('Code' = 'ICD_code')) %>%
  group_by(Dx,Code,ICD_name) %>% # Group by diagnosis, and code
  summarize(N=n()) %>% # Count every combination of diagnosis and code
  arrange(desc(N)) # Put from most common to least common

pander::pander(head(Codes_used_by_Dx, n=10))
```

This analysis by diagnosis number and code suggests that these very common codes were, as would be expected, subsidiary codes. Only 'Unspecified injury of head' 'Open wound of other parts of head' and 'Fracture of nasal bones' were common first diagnoses in this set.

```{r}
rm(Codes_used_all_Dx, Codes_used_by_Dx, Codes_used_Dx_1)
```


# Working file

## Pivot TBI to long -  one row per non-missing diagnosis

Each row is one diagnosis for one HIPE episode. Patients with many diagnoses have many rows.

```{r}
TBI_l <- TBI %>%
  select(ID,E_MRN,SOURCE,Hospital_name,starts_with('Dx')) %>%
  pivot_longer(cols = c(Dx_1_ICD_code:Dx_30_ICD_code,
         Dx_1_ICD_name:Dx_30_ICD_name),
               names_to=c("Dx",".value"),
               names_pattern="(Dx_\\d+)_(.*)") %>%
  filter(!is.na(ICD_code))

```

#Save the data files for further work

```{r}
saveRDS(Dictionary, file='data/Dictionary.Rds')
saveRDS(TBI, file='data/TBI.Rds')
saveRDS(TBI_l, file='data/TBI_l.Rds')

```
