---
title: "External codes"
author: "Anthony Staines & Anne O'Farrell"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
editor_options:
  chunk_output_type: console
---

# Process external codes

This file prepares the dictionary and the set of binary variables linking e-codes to the main groups used for analysis.

```{r setup, include=FALSE}
rm(list = ls())
library(lubridate)
library(gt)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE,cache = TRUE)
```

# Read in data

```{r}
TBI <- readRDS('data/TBI.Rds')
TBI_l <- readRDS('data/TBI_l.Rds')

Dictionary <- readRDS('data/Dictionary.Rds')

```

Order the dictionary correctly, and add the Row number as a variable

```{r}
Dictionary <- Dictionary %>%
  arrange(ICD_code) %>%
  mutate(Row = row_number())
```

# Work through codes in Oz paper

Pozzato I, Tate RL, Rosenkoetter U, Cameron ID. Epidemiology of hospitalised traumatic brain injury in the state of New South Wales, Australia: a population-based study. Australian and New Zealand Journal of Public Health [Internet]. 2019 [cited 2021 Feb 2];43(4):382–8. Available from: https://onlinelibrary.wiley.com/doi/abs/10.1111/1753-6405.12878

## Defining features of injury

The first level of codes includes those reported as a skull fracture, an intra-cranial injury, or a crush injury to the head.

### TBI Diagnosis Codes 1

Skull fracture
 Fracture of vault or base of the skull S02.0 - S02.1
 Fractures of skull and facial bones of other or unspecified nature S02.7- S02.9

Intracranial injury
 Concussion  S06.00
 Traumatic cerebral oedema S06.1
 Diffuse cerebral/cerebellar contusions or haematomas  S06.20-S06.28
 Focal cerebral/cerebellar contusion or haematoma  S06.30- S06.38
 Epidural/subarachnoid/subdural haemorrhage  S06.4- S06.6
 Intracranial injuries of other or unspecified nature  S06.8- S06.9

Crush injury of the head
 Crushing injury of skull  S07.1
 
```{r}
#
# Skull fracture
# Fracture of vault or base of the skull S02.0- S02.1
# Fractures of skull and facial bones of other or unspecified nature S02.7- S02.9
#
Dictionary %>% filter(str_detect(ICD_code, 'S02'))
Dictionary <- Dictionary %>%
  mutate(Skull = str_detect(ICD_code, 'S020|S021|S027|S028|S029')) %>%
    mutate(Skull = ifelse(Skull, Skull, NA))
    table(Dictionary$Skull)

#Intracranial injury
# Concussion  S06.00
# Traumatic cerebral oedema S06.1
# Diffuse cerebral/cerebellar contusions or haematomas  S06.20-S06.28
# Focal cerebral/cerebellar contusion or haematoma  S06.30- S06.38
# Epidural/subarachnoid/subdural haemorrhage  S06.4- S06.6
# Intracranial injuries of other or unspecified nature  S06.8- S06.9
#
Dictionary %>% filter(str_detect(ICD_code, 'S06'))
Dictionary <- Dictionary %>%
  mutate(IC_injury = str_detect(ICD_code, 'S060|S061|S062|S063|S064|S065|S066|S068|S069')) %>%
    mutate(IC_injury = ifelse(IC_injury, IC_injury, NA))
    table(Dictionary$IC_injury)

#Crush injury of the head
# Crushing injury of skull  S07.1 NONE RECORDED
# 
Dictionary %>% filter(str_detect(ICD_code, 'S07'))
Dictionary <- Dictionary %>%
  mutate(Crush = str_detect(ICD_code, 'S071')) %>%
    mutate(Crush = ifelse(Crush, Crush, NA))
    table(Dictionary$Crush)

```


### TBI diagnosis Codes #2

Loss of consciousness

 LoC of unspecified duration  S06.01
 LoC < 30 min                 S06.02
 LoC 30 min to 24 hours       S06.03
 LoC > 24 hours               S06.04-05

```{r}
#
#LOC
# LoC of unspecified duration   S06.01
# LoC < 30 min                  S06.02
# LoC 30 min to 24 hours        S06.03
# LoC > 24 hours                S06.04-05
#
Dictionary %>% filter(str_detect(ICD_code, 'S060'))
Dictionary <- Dictionary %>%
  mutate(LOC = str_detect(ICD_code, 'S0601|S0602|S0603|S0604|S0605')) %>%
    mutate(LOC = ifelse(LOC, LOC, NA)) %>%
  mutate(LOC_duration = recode(ICD_code,
                           'S0601' = 'n.s.',
                           'S0602' = '< 0.5',
                           'S0603' = '< 24',
                           'S0604' = '> 24',
                           'S0605' = '> 24',
                           .default = NA_character_
                           ))
```

## Injury-related characteristics

### Other TBI severity codes

#### Post traumatic amnesia

 PTA of unspecified duration   F04.0
 PTA < 24 hours                F04.1
 PTA 24 hours to 2 weeks       F04.2
 PTA ≥ 2 weeks                 F04.3


```{r}
#Post traumatic amnesia
# PTA of unspecified duration   F04.0
# PTA < 24 hours                F04.1
# PTA 24 hours to 2 weeks       F04.2
# PTA ≥ 2 weeks                 F04.3
#
Dictionary %>% filter(str_detect(ICD_code, 'F04'))
Dictionary <- Dictionary %>%
  mutate(PTA = str_detect(ICD_code, 'F0400|F0401|F0402|F0403')) %>%
    mutate(PTA = ifelse(PTA, PTA, NA)) %>%
  mutate(PTA_duration = recode(ICD_code,
                           'F0400' = 'n.s.',
                           'F0401' = '< 1d',
                           'F0402' = '< 2w',
                           'F0403' = '> 2w',
                           .default = NA_character_
                           ))
```

We've now got a file which classifies each code used in the data set into one of five groups Skull, IC_injury, Crush, LOC, LOC_duration, PTA, and PTA_duration.

### E-codes

External causes are a whole chapter XX in ICD-10 broken down as follows.
	XX External causes of morbidity and mortality  
  	V01-X59 Accidents  
	    V01-V99 Transport accidents
	    W00-X59 Other external causes of accidental injury
	  X60-X84 Intentional self-harm
	  X85-Y09 Assault
	  Y10-Y34 Event of undetermined intent
	  Y35-Y36 Legal intervention and operations of war
The last three blocks are not relevant
	  Y40-Y84 Complications of medical and surgical care
	  Y85-Y89 Sequelae of external causes of morbidity and mortality
	  Y90-Y98 Supplementary factors related to causes of morbidity and mortality classified elsewhere

Some subgroups of W are not relevant either
	W50-W64 Exposure to animate mechanical forces
	W65-W74 Accidental drowning and submersion
	W75-W84 Other accidental threats to breathing
	W85-W99 Exposure to electric current, radiation and extreme ambient air temperature and pressure
	X00-X09 Exposure to smoke, fire and flames
	X10-X19 Contact with heat and hot substances
	X20-X29 Contact with venomous animals and plants
	X30-X39 Exposure to forces of nature
	X40-X49 Accidental poisoning by and exposure to noxious substances
	X50-X57 Overexertion, travel and privation
	X58-X59 Accidental exposure to other and unspecified factors

#### E_block

```{r}
Dictionary %>% select(ICD_code) %>% filter(ICD_code >'U9999' & ICD_code < 'Y40')
Dictionary %>% select(ICD_code) %>% filter(ICD_code >'U9999' & ICD_code < 'Y40') %>% tail()
Dictionary %>% select(ICD_code) %>% filter(ICD_code >'W499' & ICD_code < 'X60')
Dictionary %>% select(ICD_code) %>% filter(ICD_code >'W499' & ICD_code < 'X60') %>% tail

Dictionary <- Dictionary %>%
  mutate(E_block =  ifelse((ICD_code >'U9999' & ICD_code < 'Y40') &
                             !(ICD_code >'W499' & ICD_code < 'X60'),TRUE,NA)) %>%
  mutate(ICD_group = str_extract(ICD_code, '^[A-Z]\\d{2}'))

```

#### E-codes
External cause codes (E-Codes)

Transport accidents   V01-V99
Falls                 W00-W19
Assaults              X85-Y09
Other accidents       W20-W64, X60-X84
Undetermined          Y21-Y33 # Not in Pazzato? why?

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 

```{r}
#V01 - V99 inclusive
Dictionary %>% filter(str_starts(ICD_code, 'V0')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'V99')) %>% select(ICD_code,ICD_name,Row)
Dictionary <- Dictionary %>%
  mutate(Transport = if_else(str_starts(ICD_code, 'V' ),TRUE, NA))

#W00 - W19 Falls
Dictionary %>% filter(str_starts(ICD_code, 'W0')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'W1')) %>% select(ICD_code,ICD_name,Row) %>% tail()
Dictionary <- Dictionary %>%
  mutate(Falls = ifelse ((ICD_code >= 'W0' & ICD_code <= 'W19'), TRUE, NA))
    table(Dictionary$Falls)

#X60 - X79 Intentional self harm
#X80 - X84 Intentional self harm
#
Dictionary %>% filter(str_starts(ICD_code, 'X6'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'X7'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(ICD_code > 'X8' & ICD_code < 'X85') %>% select(ICD_code,ICD_name,Row)
Dictionary <- Dictionary %>%
  mutate(Selfharm = ifelse ((ICD_code >= 'X60' & ICD_code <= 'X85'), TRUE, NA))
    table(Dictionary$Selfharm)

#X85 - Y09 Assaults
Dictionary %>% filter(str_starts(ICD_code, 'X85')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'Y09')) %>% select(ICD_code,ICD_name,Row)
Dictionary <- Dictionary %>%
  mutate(Assaults = ifelse ((ICD_code >= 'X85' & ICD_code < 'Y1'), TRUE, NA))
    table(Dictionary$Assaults)

#W20 - W64 Other accidents
Dictionary %>% filter(str_starts(ICD_code, 'W2')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'W6')) %>% select(ICD_code,ICD_name,Row)
Dictionary <- Dictionary %>%
  mutate(Other = ifelse ((ICD_code >= 'W2' & ICD_code <= 'W64'), TRUE, NA))
    table(Dictionary$Other)

# Y21 - Y33
# Undetermined
Dictionary %>% filter(str_starts(ICD_code, 'Y1')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'Y2')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'Y3')) %>% select(ICD_code,ICD_name,Row)
Dictionary <- Dictionary %>%
  mutate(Undetermined = ifelse ((ICD_code >= 'Y10' & ICD_code <= 'Y34'), TRUE, NA))
    table(Dictionary$Undetermined)

```

# Associated factors
Associated factor codes

 Drug use      F11-F16, F18-F19, X40-4, X60-4, Z72.2
 Alcohol use   F10.0-8, T51.0, T51.9, X45, X65, Y90.1-9, Z72.1   
 Sport         Y92.30-9
 Recreation    U50-U72

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 

```{r}
# Drug use
# F11-F16, F18-F19, X40-4, X60-4, Z72.2
Dictionary %>% filter(str_starts(ICD_code, 'F11')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F12')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F13')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F14')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F15')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F16')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F17')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F18')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'F19')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'X4'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'X6'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'Z72')) %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Drug = ifelse ((ICD_code >= 'F11' & ICD_code < 'F17')|
                       (ICD_code > 'F17' & ICD_code < 'F20')|
                       (ICD_code >= 'X40' & ICD_code < 'X45')|
                       (ICD_code >= 'X40' & ICD_code < 'X45')|
                       (ICD_code >= 'X60' & ICD_code < 'X65')|
                       (str_starts(ICD_code, 'Z72')), TRUE, NA))
    table(Dictionary$Drug)

#Alcohol use
# F10.0-8, T51.0, T51.9, X45, X65, Y90.1-9, Z72.1
#
Dictionary %>% filter(str_starts(ICD_code, 'F10') |  str_starts(ICD_code, 'T510') | str_starts(ICD_code, 'T519'))  %>% select(ICD_code,ICD_name,Row)

Dictionary %>% filter(str_starts(ICD_code, 'X45'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'X65'))  %>% select(ICD_code,ICD_name,Row)

Dictionary %>% filter(str_starts(ICD_code, 'Y90') | str_starts(ICD_code, 'Z721'))  %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Alcohol = ifelse(
    (str_starts(ICD_code, 'F10')) |
    (str_starts(ICD_code, 'T510')) |
    (str_starts(ICD_code, 'X45')) |
    (str_starts(ICD_code, 'X65')) |
    (str_starts(ICD_code, 'Y90')) |
    (str_starts(ICD_code, 'Z721')), TRUE, NA)) 
    table(Dictionary$Alcohol)

#Sport
#Y9230 - Y9239 Row 6408 to Row 6416
#
Dictionary %>% filter(str_starts(ICD_code, 'Y923'))  %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Sport = ifelse(str_starts(ICD_code, 'Y923'), TRUE, NA))
    table(Dictionary$Sport)

#Recreation
#U50 - U72 Row 5168 to Row 5310
Dictionary %>% filter(str_starts(ICD_code, 'U5'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'U6'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'U7'))  %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Recreation = ifelse(ICD_code >='U5000' & ICD_code < 'U7300', TRUE, NA))
    table(Dictionary$Recreation)

```

## Associated injury

 Other mechanical trauma   S10-T14
 Complications   T80-88
 Other injuries    T15-T79, T89

These have to be *manually* checked every time the data change to find the first and last (e.g. W00, W000, W009 etc...)

Checked `r Sys.Date()`  Anthony Staines 


```{r}
#Other mechanical trauma
#S10-T14
Dictionary %>% filter(str_starts(ICD_code, 'S10'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'T14'))  %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Other_mechanical = ifelse(ICD_code >= 'S100' & ICD_code < 'T15', TRUE, NA))
    table(Dictionary$Other_mechanical)

#Complications
#T80-88

Dictionary %>% filter(str_starts(ICD_code, 'T80')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'T88')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'T89')) %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Complications = ifelse(ICD_code >= 'T8' & ICD_code < 'T89', TRUE, NA))
    table(Dictionary$Complications)
  

#Other injuries
#T15-T79, T89
Dictionary %>% filter(str_starts(ICD_code, 'T15')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'T79')) %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'T89')) %>% select(ICD_code,ICD_name,Row)

Dictionary <- Dictionary %>%
  mutate(Other_injuries = ifelse((ICD_code >= 'T15' & ICD_code < 'T80')| str_starts(ICD_code, 'T89'), TRUE, NA))
    table(Dictionary$Other_injuries)

```


## Exclusions

Syncope and collapse R55

```{r}
#
# Syncope and collapse R55
#
Dictionary %>% filter(str_starts(ICD_code, 'R5'))  %>% select(ICD_code,ICD_name,Row)
Dictionary %>% filter(str_starts(ICD_code, 'R5'))  %>% select(ICD_code,ICD_name,Row) %>% tail(12)

Dictionary <- Dictionary %>%
  mutate(Syncope = ifelse(ICD_code == 'R55', TRUE, NA))
    table(Dictionary$Syncope)

```

## E_block and External

E_block is an ICD code in the relevant set of codes
External is a code in Transport, Falls, Assaults, Other

```{r}
D <- Dictionary %>%
  select(ICD_group, ICD_code, E_block,Transport, Falls, Assaults, Other, Undetermined, Selfharm) %>%
  mutate(External = ifelse((Transport | Falls | Assaults | Other), TRUE, NA))

table(D$E_block, D$External, useNA = 'ifany')

rm(D)
```


Following a manual check all the codes in the E_Block with no External code should not have an External code


#Merge the codes
Join the external code groups to the individual files
 E_block, ICD_group
 Transport, Falls, Assaults, Other
 Selfharm, Undetermined,
 Drug, Alcohol, Sport, Recreation
 Other_mechanical, Complications, Other_injuries
 Syncope

```{r}
names(TBI_l)

TBI_l <- TBI_l %>%
  left_join(Dictionary, by = c("ICD_code", "ICD_name"))

names(TBI_l)

```

# Widen the file
The main file TBI_l now has one row per diagnosis. For many purposes we will need one row per person.

```{r}
#The subfiles come in numerical order
#

#TBI_0
# This file has the ID, the E_MRN, and all the diagnoses in a row.
#
TBI_0 <- TBI_l %>%
  pivot_wider(id_cols = c(ID, E_MRN, SOURCE), names_from = Dx, values_from = c(ICD_code, ICD_name))

#TBI_1
# Convert the per-diagnosis classifications of ancillary variables to per-person
# Each is a group of up to 30 values, TRUE or NA
# ~as.logical(sum(!is.na(.)))
# The missing values are omitted
# The non-missing values are summed
# The total is converted to a logical value 0  = False, 1 or more = TRUE
#
TBI_1 <- TBI_l %>%
  select(ID,E_MRN,Skull:LOC,PTA,E_block, Transport:Syncope) %>% # Leave out durations and ICD_group
  group_by(ID, E_MRN) %>%
  summarise(across(Skull:Syncope, ~as.logical(sum(!is.na(.))))) # Total of non-missing values as a 0,1 logical

# Quick check on codes for LOC and PTA duration
unique(TBI_l$LOC_duration)
unique(TBI_l$PTA_duration)

#TBI_2
# There are two duration variables, one for LOC and one for PTA
# Some people have more than one, which is surprising
#
TBI_2 <- TBI_l %>%
  select(ID,E_MRN,LOC_duration,PTA_duration) %>%
  group_by(ID, E_MRN) %>%
  mutate(LOCl = ifelse(is.na(LOC_duration),NA,TRUE)) %>%
  mutate(PTAl = ifelse(is.na(PTA_duration),NA,TRUE))

table(TBI_2$LOC_duration, useNA = 'ifany')
table(TBI_2$PTA_duration, useNA = 'ifany')
table(TBI_2$LOC_duration, TBI_2$PTA_duration, useNA = 'ifany')

# None have two PTA codes
Check <- TBI_2 %>%
  group_by(ID,E_MRN) %>%
  summarise(across(LOCl:PTAl, ~sum(!is.na(.)))) %>%
  filter(PTAl > 1)

Check$ID 
Check$E_MRN
#
#Three have two LOC codes provided
Check <- TBI_2 %>%
  group_by(ID,E_MRN) %>%
  summarise(across(LOCl:PTAl, ~sum(!is.na(.)))) %>%
  filter(LOCl > 1)

Check$ID
Check$E_MRN

#One row each - these are not repeated records on the same people
TBI %>% filter(E_MRN == 'CVEATIY') %>% summarise(N=n()) #1
TBI %>% filter(E_MRN == 'COE5UI3') %>% summarise(N=n()) #1
TBI %>% filter(E_MRN == '8O#3VF1') %>% summarise(N=n()) #1

#
#Manual fix
# CVEATIY Two codes, S0602 and S0601 brief and unspecified, Dx1 and Dx5, Choose brief
# COE5UI3 Two codes, S0601 and S0602 unspecified and brief, Dx2 and Dx15, Choose unspecified
# 8O#3VF1 Two codes, S0601 and S0602, brief and unspecified, Dx5 and Dx6, Choose brief
#
#
# Print out codes
#
TBI_0 %>%
  filter(E_MRN == 'CVEATIY') %>% paste()
TBI_0 %>%
  filter(E_MRN == 'COE5UI3') %>% paste()
TBI_0 %>%
  filter(E_MRN == '8O#3VF1') %>% paste()

#
#Set the surplus LOC codes to NA
#
TBI_2$LOC_duration[TBI_2$E_MRN == 'CVEATIY' & TBI_2$LOC_duration == 'n.s.'] = NA_character_
TBI_2$LOC_duration[TBI_2$E_MRN == 'COE5UI3' & TBI_2$LOC_duration == '< 0.5'] = NA_character_
TBI_2$LOC_duration[TBI_2$E_MRN == '8O#3VF1' & TBI_2$LOC_duration == 'n.s.'] = NA_character_

#Check
# Recalculate logical variables with the new data
#
TBI_2 <- TBI_2 %>%
  group_by(ID, E_MRN) %>%
  mutate(LOCl = ifelse(is.na(LOC_duration),NA,TRUE)) %>%
  mutate(PTAl = ifelse(is.na(PTA_duration),NA,TRUE))

#Three had two LOC codes provided
#Should be fixed by now
#
Check <- TBI_2 %>%
  group_by(ID,E_MRN) %>%
  summarise(across(LOCl:PTAl, ~sum(!is.na(.)))) %>%
  filter(LOCl > 1)

Check$ID 
#Now zero length
#

# Recode the LOC and PTA durations as numbers
#
TBI_2 <- TBI_2 %>%
  mutate(LOCn = recode(LOC_duration,
                       'n.s.' = 0,
                       '< 0.5' = 1,
                       '< 24' = 2,
                       '> 24' = 3,
                       )) %>%
  mutate(PTAn = recode(PTA_duration,
                       'n.s.' = 0,
                       '< 1d' = 1,
                       '< 2w' = 2,
                       '> 2w' = 3,
                       ))

#
#TBI_3
#  Cuts down to files with LOC
#
TBI_3 <- TBI_2 %>%
  select(ID,E_MRN, LOC_duration, LOCl, LOCn) %>%
  filter(!is.na(LOC_duration) | !is.na(LOCl) | !is.na(LOCn))

#
#More than one?
#
Check <- TBI_3 %>%
  select(ID,E_MRN) %>%
  group_by(ID,E_MRN) %>%
  summarise(N = n()) %>%
  filter(N != 1)
Check
#All zero

#
#TBI_4
#  Cuts down to files with PTA
#
TBI_4 <- TBI_2 %>%
  select(ID,E_MRN, PTA_duration, PTAl, PTAn) %>%
  filter(!is.na(PTA_duration) | !is.na(PTAl) | !is.na(PTAn))

#
#More than one?
#
Check <- TBI_4 %>%
  select(ID,E_MRN) %>%
  group_by(ID,E_MRN) %>%
  summarise(N = n()) %>%
  filter(N != 1)
Check
#All zero

#
#Additional variables from TBI
#
TBI_5 <- TBI %>%
  select(ID,E_MRN,Hospital_name:Team_code_principal,Transfer_from:Dx_1_name_CCS_IM,Proc_1_ACHI_name:Age_group)


#
#Merge the files
# Diagnoses TBI_0,
# Skull:Syncope TB_1,
# LOC TBI_3
# PTA TBI_4
# Admission level variables TBI_5

TBI_w <- TBI_1 %>%
  left_join(TBI_3, by = c("ID", "E_MRN")) %>%
  left_join(TBI_4, by = c("ID", "E_MRN")) %>%
  left_join(TBI_5, by = c("ID", "E_MRN")) %>%
  left_join(TBI_0, by = c("ID", "E_MRN")) %>% # Puts diagnoses at the end
  ungroup()

names(TBI_w)
table(TBI_w$SOURCE)

rm(TBI_0, TBI_1, TBI_2, TBI_3, TBI_4, TBI_5, Check)
```

# Code multiple E-causes

A small number of records have more than one external cause.

```{r}

TBI_w <- TBI_w %>%
  mutate(External = ifelse((Transport + Falls + Assaults + Other > 0), TRUE, FALSE)) %>% #Any E-code
  mutate(Multiple = ifelse((Transport + Falls + Assaults + Other > 1), TRUE, FALSE)) %>%
  mutate(Transport_only = ifelse(Multiple,FALSE,Transport)) %>%
  mutate(Falls_only     = ifelse(Multiple,FALSE,Falls    )) %>%
  mutate(Assaults_only  = ifelse(Multiple,FALSE,Assaults )) %>%
  mutate(Other_only     = ifelse(Multiple,FALSE,Other    )) # Other, but not Multiple...

#Check
table(TBI_w$External)
table(TBI_w$External, TBI_w$E_block)

table(TBI_w$Multiple)
table(TBI_w$Transport,TBI_w$Multiple)
table(TBI_w$Transport_only,TBI_w$Multiple)
table(TBI_w$Falls,TBI_w$Multiple)
table(TBI_w$Falls_only,TBI_w$Multiple)
table(TBI_w$Assaults,TBI_w$Multiple)
table(TBI_w$Assaults_only,TBI_w$Multiple)
table(TBI_w$Other,TBI_w$Multiple)
table(TBI_w$Other_only,TBI_w$Multiple)

TBI_w <- TBI_w %>%
  mutate(E_group = if_else(External,'***',NA_character_)) %>%
  mutate(E_group = if_else(Transport_only,'Transport',E_group)) %>%
  mutate(E_group = if_else(Falls_only,'Falls',E_group)) %>%
  mutate(E_group = if_else(Assaults_only,'Assaults',E_group)) %>%
  mutate(E_group = if_else(Other_only,'Other',E_group)) %>%
  mutate(E_group = if_else(Multiple,'Multiple',E_group))
  
  
table(TBI_w$External, useNA = 'ifany')
table(TBI_w$E_group, useNA = 'ifany')
table(TBI_w$E_block, useNA = 'ifany')
table(TBI_w$External,TBI_w$E_group, useNA = 'ifany')
table(TBI_w$E_block,TBI_w$E_group, useNA = 'ifany')

Check1 <- TBI_w %>% filter(!E_block & External)

table(TBI_w$Gender,TBI_w$E_group, useNA = 'ifany')
table(TBI_w$Age_group,TBI_w$E_group, useNA = 'ifany')
table(TBI_w$External,TBI_w$Undetermined, useNA = 'ifany')
table(TBI_w$E_group,TBI_w$Undetermined, useNA = 'ifany')

```

## Check to see if there are multiple intents

```{r}

table(TBI_w$Selfharm)
table(TBI_w$Undetermined)
table(TBI_w$Selfharm, TBI_w$Undetermined)

TT <- TBI_w %>% filter(Selfharm & Undetermined) # All five are likely self harm, based on the codes given
TT$ID
#  1034  7154 15230 26834 42655
TBI_w <- TBI_w %>%
  mutate(Undetermined = ifelse(Selfharm & Undetermined, FALSE, Undetermined))

table(TBI_w$Selfharm, TBI_w$Undetermined)
rm(TT)
```

# Check to see matching worked ok

```{r}
Original <- TBI %>% select(ID, E_MRN, SOURCE, Hospital_name, Admission_date, Discharge_date, Team_code_admission, Team_code_principal) 
Coded    <- TBI_w %>% select(ID, E_MRN, SOURCE, Hospital_name, Admission_date, Discharge_date, Team_code_admission, Team_code_principal) 

Original %>% anti_join(Coded)
Coded %>% anti_join(Original)
```
It did.

#Save the data files for further work

```{r}
saveRDS(TBI_w, file='data/TBI_w.Rds')

```

