---
title: "HIPE analysis"
author: "Anthony Staines & Anne O\\'Farrell"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
library(csodata)
library(epitools)
library(lubridate)
library(gt)
library(kableExtra)
library(tidyverse)
library(arsenal)

knitr::opts_chunk$set(echo = TRUE,cache = TRUE, message= FALSE, results = 'hide')
```

# Read in data

```{r}
TBI_final <- readRDS('data/TBI_final.Rds')

Dictionary <- readRDS('data/Dictionary.Rds')

```

Reorder the E_group by frequency of occurrence
 
```{r}

TBI_final <- TBI_final %>% mutate(E_group.f = fct_infreq(E_group))

```

Get the population

```{r, echo = FALSE}
Age_LIST <- c('Omit' = 'All ages',
'0-4' = 'Under 1 year',
'0-4' = '1 year',
'0-4' = '2 years',
'0-4' = '3 years',
'0-4' = '4 years',
'5-9' = '5 years',
'5-9' = '6 years',
'5-9' = '7 years',
'5-9' = '8 years',
'5-9' = '9 years',
'10-14' = '10 years',
'10-14' = '11 years',
'10-14' = '12 years',
'10-14' = '13 years',
'10-14' = '14 years',
'15-19' = '15 years',
'15-19' = '16 years',
'15-19' = '17 years',
'15-19' = '18 years',
'15-19' = '19 years',
'20-24' = '20 years',
'20-24' = '21 years',
'20-24' = '22 years',
'20-24' = '23 years',
'20-24' = '24 years',
'25-29' = '25 years',
'25-29' = '26 years',
'25-29' = '27 years',
'25-29' = '28 years',
'25-29' = '29 years',
'30-34' = '30 years',
'30-34' = '31 years',
'30-34' = '32 years',
'30-34' = '33 years',
'30-34' = '34 years',
'35-39' = '35 years',
'35-39' = '36 years',
'35-39' = '37 years',
'35-39' = '38 years',
'35-39' = '39 years',
'40-44' = '40 years',
'40-44' = '41 years',
'40-44' = '42 years',
'40-44' = '43 years',
'40-44' = '44 years',
'45-49' = '45 years',
'45-49' = '46 years',
'45-49' = '47 years',
'45-49' = '48 years',
'45-49' = '49 years',
'50-54' = '50 years',
'50-54' = '51 years',
'50-54' = '52 years',
'50-54' = '53 years',
'50-54' = '54 years',
'55-59' = '55 years',
'55-59' = '56 years',
'55-59' = '57 years',
'55-59' = '58 years',
'55-59' = '59 years',
'60-64' = '60 years',
'60-64' = '61 years',
'60-64' = '62 years',
'60-64' = '63 years',
'60-64' = '64 years',
'65-69' = '65 years',
'65-69' = '66 years',
'65-69' = '67 years',
'65-69' = '68 years',
'65-69' = '69 years',
'70-74' = '70 years',
'70-74' = '71 years',
'70-74' = '72 years',
'70-74' = '73 years',
'70-74' = '74 years',
'75-79' = '75 years',
'75-79' = '76 years',
'75-79' = '77 years',
'75-79' = '78 years',
'75-79' = '79 years',
'80-84' = '80 years',
'80-84' = '81 years',
'80-84' = '82 years',
'80-84' = '83 years',
'80-84' = '84 years',
'85-89' = '85 years',
'85-89' = '86 years',
'85-89' = '87 years',
'85-89' = '88 years',
'85-89' = '89 years',
'90-94' = '90 years',
'90-94' = '91 years',
'90-94' = '92 years',
'90-94' = '93 years',
'90-94' = '94 years',
'95+' = '95 years',
'95+' = '96 years',
'95+' = '97 years',
'95+' = '98 years',
'95+' = '99 years and over')

```

```{r}
toc <- cso_get_toc()
PEA11 <- cso_get_data('PEA11')
PEA11 <- PEA11 %>%
  filter(Single.Year.of.Age != 'All ages') %>%
  filter(Sex != 'Both sexes') %>%
  mutate(Age_group = Single.Year.of.Age) %>%
  mutate(Age_group = fct_recode(Age_group, !!!Age_LIST)) %>%
  mutate(Age_group = fct_drop(Age_group))


table(PEA11$Single.Year.of.Age)
table(PEA11$Age_group)
table(PEA11$Age_group,PEA11$Single.Year.of.Age)
str(PEA11)
str(TBI_final$Age_group)

PEA11 <- PEA11 %>%
  select(Sex,`2016`,Age_group) %>%
  group_by(Sex,Age_group) %>% # Add up single years of age.
  summarise(Pop_2016 = sum(`2016`))

```

There are two slightly different questions to be considered. The first is to look at the use of resources for TBI. For this analysis only elective admissions need to be excluded from the number of events. The second is to look at occurrence of events, and this requires excluding elective care, as well as transfers and elective readmissions.

```{r}
# Eliminate non-emergency admissions
TBI_usage <- TBI_final %>%
  filter(Admission_type == 'Emergency')

table(TBI_usage$Detail)

write_rds(TBI_usage, file = 'data/TBI_usage.Rds')


# Eliminate transfers too
TBI_occurrence <- TBI_final %>%
  filter(Admission_type == 'Emergency' | Admission_type == 'Emergency Readmission') %>%
  filter(Admission_source != 'Transfer from Acute Hospital')

table(TBI_occurrence$Detail)

write_rds(TBI_occurrence, file = 'data/TBI_occurrence.Rds')

```

```{r}
TBI_final %>% filter(((Skull | IC_injury | PTA) & External)) %>% summarise(N=n()) # 26085
TBI_final %>% filter(((Skull | IC_injury | PTA) & !External)) %>% summarise(N=n()) # 1896

TBI_final %>% filter(Detail == 'TBI') %>% summarise(N=n()) # 321
TBI_final %>% filter(((Skull | IC_injury | PTA) & External) | Detail == 'TBI') %>% summarise(N=n()) # 26406

CIs_final <-round(100*Hmisc::binconf(321,26406),2)

CIs_final = 100 - CIs_final
```

```{r}
TBI_usage %>% filter(Detail == 'TBI') %>% summarise(N=n()) # 290
TBI_usage %>% filter(((Skull | IC_injury | PTA) & External) | Detail == 'TBI') %>% summarise(N=n()) # 24987

CIs_usage <- round(100*Hmisc::binconf(290,24987),2)

CIs_usage = 100 - CIs_usage
```

```{r}
TBI_occurrence %>% filter(Detail == 'TBI') %>% summarise(N=n()) # 214
TBI_occurrence %>% filter(((Skull | IC_injury | PTA) & External) | Detail == 'TBI') %>% summarise(N=n()) # 22604

CIs_occurrence <- round(100*Hmisc::binconf(214,22604),2)

CIs_occurrence = 100 - CIs_occurrence

CIS <- cbind(c('All cases','Resource usage','Occurrence'),
             rbind(CIs_final,CIs_usage,CIs_occurrence))
str(CIS)
dimnames(CIS)[[2]][1] = 'Group'
```

For the whole set of TBI cases, as well as for the two subsets considered, the Possato et al. method seems to have excellent sensitivity.

```{r, results ='asis'}
pander::pander(CIS, caption='Estimated sensitivity and 95% confidence intervals.')
```



# Figure 1
Stacked bar chart by age and cause
This graphic shows the ways in which major change happnes by age and gender in the pattern of causes of TBI.

```{r}
ggCauses <- ggplot(data = TBI_occurrence %>% filter(!is.na(E_group)),
       aes(x = Age_group,
           group = E_group.f,
           fill = E_group.f,
           colour=E_group.f)) +
  geom_bar(position = 'fill', colour='darkblue') +
  facet_wrap(~Gender, nrow = 1) +
  ggtitle('Main external causes of injury by age') +
  scale_y_continuous(labels =
                       scales::percent_format(scale = 100)) +
  scale_fill_brewer(type='qual', palette= 'Paired',
                    guide_legend(title = 'External\ncause')) +
  scale_x_discrete() +
  theme(axis.text.x = element_text(angle = -90))

ggCauses

ggsave(plot=ggCauses, filename='images/Fig1_Causes_by_Age_Sex.png', width=8, height=4, dpi=1200)
```

# Figure 2
Rates by age and sex

```{r}
Years = 7 + 8/12 # 2013 to 2019 and first 8 months of 2020.
TT <- TBI_occurrence %>%
  group_by(Gender,Age_group) %>%
  summarise(N = n()) %>%
  ungroup() %>%
  left_join(PEA11, by=c('Age_group' = 'Age_group','Gender' = 'Sex')) 

TT <- TT %>%
  mutate(Rate = 100000 * N/Pop_2016) %>%
  mutate(LrCI = 100000 * pois.daly(TT$N,TT$Pop_2016)$lower) %>%
  mutate(UprCI = 100000 * pois.daly(TT$N,TT$Pop_2016)$upper) 


ggFig2 <- ggplot(TT,aes(x=Age_group, y=Rate, group=Gender, colour=Gender)) +
  geom_ribbon(aes(y=Rate, ymin=LrCI, ymax=UprCI),colour='lightgrey', fill='lightgrey', alpha = 0.8) +
  geom_line() +
  labs(title = 'Rate of TBI by gender',
       subtitle = 'Approximate 95% confidence intervals',
       x = 'Age group',
       y = 'Rate per 100,000 per year',
       caption='Reported rates of TBI by age and sex - source HIPE data') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = -90))

ggFig2

ggsave(plot=ggFig2, filename='images/Fig2_Age_Sex_Rate.png', width=8, height=4, dpi=1200)

```

As expected the overall rate of TBI is higher in men than women, and rises with age. There is a distinct excess in young men from 15 to 24.

## Rates by age and sex and year

```{r}
TTa <- TBI_occurrence %>%
  group_by(Admission_year, Gender,Age_group) %>%
  summarise(N = n()) %>%
  ungroup() %>%
  left_join(PEA11, by=c('Age_group' = 'Age_group','Gender' = 'Sex')) 

TTa <- TTa %>%
  mutate(Rate = 100000 * N/Pop_2016) %>%
  mutate(LrCI = 100000 * pois.daly(TTa$N,TTa$Pop_2016)$lower) %>%
  mutate(UprCI = 100000 * pois.daly(TTa$N,TTa$Pop_2016)$upper) 


ggFig2a <- ggplot(TTa,aes(x=Age_group, y=Rate, group=Gender, colour=Gender)) +
  geom_ribbon(aes(y=Rate, ymin=LrCI, ymax=UprCI),colour='lightgrey', fill='lightgrey', alpha = 0.8) +
  geom_line() +
  labs(title = 'Rate of TBI by gender',
       subtitle = 'Approximate 95% confidence intervals',
       x = 'Age group',
       y = 'Rate per 100,000 per year',
       caption='Changes in reported rates of TBI by age and sex from 2013 to 2020 - source HIPE data') +
  coord_cartesian(ylim = c(0,1250)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = -90)) +
  facet_wrap(~Admission_year)

ggFig2a

ggsave(plot=ggFig2a, filename='images/Fig2a_Age_Sex_Rate_by_Year.png', width=8, height=4, dpi=1200)

```

This shows a similar pattern for each year, but the reported incidence rises steadily over time in the older age groups.

```{r}

TTb <- TTa %>% inner_join(TT,  by = c("Gender", "Age_group")) %>%
  mutate(RR = (N.x/Pop_2016.x)/(N.y/Pop_2016.y))

```

On further analysis


# Preparatory tables

```{r}
Tab1 <- tableby( ~ Admission_type + Admission_source, data=TBI_occurrence)

summary(Tab1, text = TRUE)
```

# Table 1

```{r, results="asis"}
TBI_occurrence <- TBI_occurrence %>%
  mutate(DrugUse = as.numeric(Drug)) %>%
  mutate(AlcoholUse = as.numeric(Alcohol))
  
CONTROL <- tableby.control(numeric.stats = 'meansd')

Tab1 <- tableby(Gender ~ Age_by_1_year + LOS_total + ITU_or_CCU_bed_days + DrugUse + AlcoholUse, data=TBI_occurrence, control = CONTROL)

summary(Tab1, text = TRUE)
```

```{r}
Daily_Admissions <- TBI_occurrence %>%
  select(Admission_date) %>%
  mutate(Day_of_week = wday(Admission_date,
                            label = TRUE))  %>%
  mutate(Weekend = ifelse((Day_of_week %in% c('Sat','Sun')), TRUE, FALSE))

Daily_Admissions %>%
  group_by(Day_of_week) %>%
  summarise(N = n())

Daily_Admissions %>%
  summarise(N = n())

ggplot(Daily_Admissions %>%
         group_by(Weekend, Day_of_week, Admission_date) %>%
         summarise(N = n()),
       aes(x=Admission_date, y=N)) +
  geom_point(alpha=0.7, aes(colour=Day_of_week, fill = Day_of_week)) +
  geom_smooth( method = "gam", formula = y ~ s(x, k=20, bs = "cs"))  +
#  geom_line(alpha=0.25,colour='darkblue') +
  scale_x_date(date_minor_breaks='3 months',
               date_breaks = '1 year',
               date_labels = '%b-%Y',
               limits = c(as_date('2013-01-01'),NA)) +
  scale_colour_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  ggtitle('Daily admissions 2013-2020', subtitle = 'HIPE data') +
  xlab('Date of admission') + ylab('Number of admissions')

ggplot(Daily_Admissions, aes(x=Day_of_week)) +
  geom_bar(aes(fill=Day_of_week)) +
  scale_fill_viridis_d() +
  theme_minimal()

rm(Daily_Admissions)
```
